---
title: "Effects of multiple disturbance types"
author: "Kai Kopecky"
date: "2/12/2021"
output: html_document
---
### The model
#### State Variables for proportional cover of benthic space holders
$$
\begin{align}
&C_L\ =\ Live\ coral\ cover\\
\newline
&C_D\ =\ Dead\ coral\ cover\\
\newline
&M_S\ =\ Macroalgae\ cover\ on\ open\ substrate\\
\newline
&M_C\ =\ Macroalgae\ cover\ on\ dead\ coral\\
\newline
&S\ =\ Open\ Substrate\\
\end{align}
$$

#### Ordinary differential equations 
Rates of change for proportional spatial cover of live coral, dead coral, and macroalgae on either open substrate or dead coral
$$
\begin{align}
&\frac{d C_L}{dt} =  g_c C_L S + \gamma_cS - g_{mC_L} C_L(M_S + M_C) - d_c C_L \\ 
\newline
&\frac{d C_D}{dt} = d_c C_L + M_C(h_{mc} + d_m) - g_m C_D (M_S + M_C) - e C_D \\ 
\newline
&\frac{d M_S}{dt} = g_m M_S S + g_m M_C S + \gamma_mS - M_S(h_{ms} + d_m)\\
\newline
&\frac{d M_C}{dt} = g_m M_C C_D +  g_m M_S C_D + g_{mC_L} C_L(M_S + M_C) + \gamma_mC_D - M_C(h_{mc} + d_m + e)\\
\newline
&\ S = 1 - C_L - C_D - M_S - M_C \\ 
\newline
\end{align}
$$

#### Parameters for ODE's
$$
\begin{align}
&g_c\ =\ Growth\ rate\ of\ live\ coral\\
&g_m\ =\ Growth\ rate\ of\ macroalgae\\
&g_{mC_L}\ =\ Growth\ rate\ of\ macroalgae\ onto\ live\ coral\\
&d_c\ =\ Ambient\ mortality\ of\ coral\\
&d_m\ =\ Ambient\ mortality\ of\ macroalgae\\
&e\ =\ Erosion\ of\ dead\ coral\\
&\gamma_c\ =\ Settlement\ rate\ of\ coral\\
&\gamma_m\ =\ Settlement\ rate\ of\ macroalgae\\
&h_{ms}\ =\ herbivory\ rate\ of\ macroalgae\ on\ open\ substrate\\
&h_{mc}\ =\ herbivory\ rate\ of\ macroalgae\ on\ dead\ coral\\
&\end{align}
$$

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Parameter values (per year), echo = FALSE}
# Set parameter values
g_C <- 0.15
d_C <- 0.002
gamma_C <- 0.001
g_M <- 0.5
g_M_C_L <- 0.025
d_M <- 0.05
h_M_S <- 0.25
h_M_C <- 0.2 # default value
#h_M_C.two_thirds <- 0.2
#h_M_C.half <- 0.15
#h_M_C.third <- 0.1
gamma_M <- 0.0001
e <- 0.01
#t_disturb <- 200 # time at which single disturbance occurs

t_disturb_1 <- 120 # disturbance times
t_disturb_2 <- 180
t_disturb_3 <- 240
t_disturb_4 <- 300
t_disturb_5 <- 360
t_disturb_6 <- 420
t_disturb_7 <- 480
t_disturb_8 <- 540
t_disturb_9 <- 600
t_disturb_10 <- 660
t_disturb_11 <- 720
t_disturb_12 <- 780
t_disturb_13 <- 840
t_disturb_14 <- 900
t_disturb_15 <- 960
t_disturb_16 <- 1020
t_disturb_17 <- 1080
t_disturb_18 <- 1140
#disturbance_times.ten_years <- seq(from = 120, to = 1080, by = 120)
#disturbance_times.twenty_years <- seq(from = 100, to = 1200, by = 240)

disturb_set <- seq(from = 0, to = 1, by = .01) # Time vector for outer loop
tset.disturb <- seq(from = 0, to = 1200) # Time vector for inner loop

```

### Multiple disturbances

#### Time series

50% removal of all space holders every ten years
```{r 50% structure-removing disturbance ten year frequency, echo = FALSE}

# Create time vector, holding vectors for each state variable, and set initial conditions
tset_phys_dist <- seq(from = 0, to = 1200)
C_L.simu_phys <- NaN*tset_phys_dist; C_L.simu_phys[1] <- 0.6
C_D.simu_phys <- NaN*tset_phys_dist; C_D.simu_phys[1] <- 0
M_S.simu_phys <- NaN*tset_phys_dist; M_S.simu_phys[1] <- 0.05
M_C.simu_phys <- NaN*tset_phys_dist; M_C.simu_phys[1] <- 0.05

# For() loop with if/else statement to simulate trajectories of spatial occupancy by live coral and macroalgae before and after disturbance

for(i in 2:length(tset_phys_dist)){ # For every element after the first in the time vector...
  
  if(tset_phys_dist[i] !=t_disturb_1 & tset_phys_dist[i] !=t_disturb_2 & tset_phys_dist[i] !=t_disturb_3 & tset_phys_dist[i] !=t_disturb_4 & tset_phys_dist[i] !=t_disturb_5 & tset_phys_dist[i] !=t_disturb_6 & tset_phys_dist[i] !=t_disturb_7 & tset_phys_dist[i] !=t_disturb_8 & tset_phys_dist[i] !=t_disturb_9){ # ...if the time point does not equal the time of disturbance, run the simulation as normal...
  
  dt <- tset_phys_dist[i] - tset_phys_dist[i-1] # calculate change in time  
 
  # Proportional cover at previous time step  
  C_L <- C_L.simu_phys[i-1]
  C_D <- C_D.simu_phys[i-1]
  M_S <- M_S.simu_phys[i-1]
  M_C <- M_C.simu_phys[i-1]
  
  # Change in proportional cover, converted to units of months (/5)
  dC_L <- (dt/5)*(g_C*C_L*(1-C_L-C_D-M_S-M_C) + gamma_C*(1-C_L-C_D-M_S-M_C) - g_M_C_L*C_L*(M_S + M_C) - d_C*C_L) 
  dC_D <- (dt/5)*(d_C*C_L + M_C*(h_M_C + d_M) - g_M*C_D*(M_S + M_C) - e*C_D)
  dM_S <- (dt/5)*(g_M*M_S*(1-C_L-C_D-M_S-M_C) + g_M*M_C*(1-C_L-C_D-M_S-M_C) + gamma_M*(1-C_L-C_D-M_S-M_C) - M_S*(h_M_S + d_M))
  dM_C <- (dt/5)*(g_M*M_C*C_D + g_M*M_S*C_D + g_M_C_L*C_L*(M_S + M_C) + gamma_M*C_D - M_C*(h_M_C + d_M + e))
 
  # Proportional cover at new time step   
  C_L.simu_phys[i] <- C_L.simu_phys[i-1] + dC_L
  C_D.simu_phys[i] <- C_D.simu_phys[i-1] + dC_D
  M_S.simu_phys[i] <- M_S.simu_phys[i-1] + dM_S
  M_C.simu_phys[i] <- M_C.simu_phys[i-1] + dM_C
  }
  
    else { # ... but if the time step does equal the time of disturbance...
  
  dt <- tset_phys_dist[i] - tset_phys_dist[i-1]

  # reduce all space holders to 50% of their pre-disturbance levels
  C_L <- C_L.simu_phys[i-1]*0.5
  C_D <- C_D.simu_phys[i-1]*0.5
  M_S <- M_S.simu_phys[i-1]*0.5
  M_C <- M_C.simu_phys[i-1]*0.5
  
  dC_L <- (dt/5)*(g_C*C_L*(1-C_L-C_D-M_S-M_C) + gamma_C*(1-C_L-C_D-M_S-M_C) - g_M_C_L*C_L*(M_S + M_C) - d_C*C_L) 
  dC_D <- (dt/5)*(d_C*C_L + M_C*(h_M_C + d_M) - g_M*C_D*(M_S + M_C) - e*C_D)
  dM_S <- (dt/5)*(g_M*M_S*(1-C_L-C_D-M_S-M_C) + g_M*M_C*(1-C_L-C_D-M_S-M_C) + gamma_M*(1-C_L-C_D-M_S-M_C) - M_S*(h_M_S + d_M))
  dM_C <- (dt/5)*(g_M*M_C*C_D + g_M*M_S*C_D + g_M_C_L*C_L*(M_S + M_C) + gamma_M*C_D - M_C*(h_M_C + d_M + e))
  
  C_L.simu_phys[i] <- C_L+ dC_L
  C_D.simu_phys[i] <- C_D+ dC_D
  M_S.simu_phys[i] <- M_S+ dM_S
  M_C.simu_phys[i] <- M_C+ dM_C
  }
  
}

# Plot of time series
plot(tset_phys_dist, C_L.simu_phys, col = "blue", lwd = 1, lty = 'solid', type = 'l', xlab = "Time (months)", ylab = "Proportional cover", xlim = c(0, max(tset_phys_dist)), ylim = c(0, 1))
lines(tset_phys_dist, C_D.simu_phys, col = "black", lwd = 1, lty = 'solid')
lines(tset_phys_dist, M_S.simu_phys, col = "green", lwd = 1, lty = 'solid')
lines(tset_phys_dist, M_C.simu_phys, col = "grey", lwd = 1, lty = 'solid')
#legend(x=350,y=0.7,legend=c('Live coral','Dead coral','Macroalgae (open substrate)','Macroalgae (dead coral)'),lwd=1,lty='solid',col=c('blue','black','green','grey'))
#abline(v=100, lty = 'dashed')
#abline(v=220, lty = 'dashed')
#abline(v=340, lty = 'dashed')
#abline(v=460, lty = 'dashed')
```

Structure-retaining disturbances: 50% reduction in live coral cover that is transferred to dead coral every ten years; macroalgae unaffected 
```{r 50% structure-retaining ten year frequency , echo = FALSE}

# Create time vector, holding vectors for each state variable, and set initial conditions
tset_bio_dist <- seq(from = 0, to = 1200)
C_L.bio <- NaN*tset_bio_dist; C_L.bio[1] <- 0.6
C_D.bio <- NaN*tset_bio_dist; C_D.bio[1] <- 0
M_S.bio <- NaN*tset_bio_dist; M_S.bio[1] <- 0.05
M_C.bio <- NaN*tset_bio_dist; M_C.bio[1] <- 0.05

# For() loop with if/else statement to simulate trajectories of spatial occupancy by live coral and macroalgae before and after disturbance
for(i in 2:length(tset_bio_dist)){ # For every element after the first in the time vector...
  
 if(tset_bio_dist[i] !=t_disturb_1 & tset_bio_dist[i] !=t_disturb_2 & tset_bio_dist[i] !=t_disturb_3 & tset_bio_dist[i] !=t_disturb_4 & tset_bio_dist[i] !=t_disturb_5 & tset_bio_dist[i] !=t_disturb_6 & tset_bio_dist[i] !=t_disturb_7 & tset_bio_dist[i] !=t_disturb_8 & tset_bio_dist[i] !=t_disturb_9){ # ...if the time point does not equal the time of disturbance, run the simulation as normal...

# calculate change in time  
  dt <- tset_bio_dist[i] - tset_bio_dist[i-1] 

  # Proportional cover at previous time step  
  C_L <- C_L.bio[i-1] 
  C_D <- C_D.bio[i-1]
  M_S <- M_S.bio[i-1]
  M_C <- M_C.bio[i-1]
 
# Change in proportional cover, converted to units of months (/5)
  dC_L <- (dt/5)*(g_C*C_L*(1-C_L-C_D-M_S-M_C) + gamma_C*(1-C_L-C_D-M_S-M_C) - g_M_C_L*C_L*(M_S + M_C) - d_C*C_L) 
  dC_D <- (dt/5)*(d_C*C_L + M_C*(h_M_C + d_M) - g_M*C_D*(M_S + M_C) - e*C_D)
  dM_S <- (dt/5)*(g_M*M_S*(1-C_L-C_D-M_S-M_C) + g_M*M_C*(1-C_L-C_D-M_S-M_C) + gamma_M*(1-C_L-C_D-M_S-M_C) - M_S*(h_M_S + d_M))
  dM_C <- (dt/5)*(g_M*M_C*C_D + g_M*M_S*C_D + g_M_C_L*C_L*(M_S + M_C) + gamma_M*C_D - M_C*(h_M_C + d_M + e))

# Proportional cover at new time step  
  C_L.bio[i] <- C_L.bio[i-1] + dC_L
  C_D.bio[i] <- C_D.bio[i-1] + dC_D
  M_S.bio[i] <- M_S.bio[i-1] + dM_S
  M_C.bio[i] <- M_C.bio[i-1] + dM_C
  
  }

  # ... but if the time step does equal the time of disturbance...
    else {
  
  dt <- tset_bio_dist[i] - tset_bio_dist[i-1]

  C_L <- C_L.bio[i-1]*0.5 # reduce live coral to 50% of its pre-disturbance level, and transfer this directly to the dead coral class
  C_D <- C_D.bio[i-1] + (C_L.bio[i-1]-C_L)
  M_S <- M_S.bio[i-1]
  M_C <- M_C.bio[i-1]
  
  dC_L <- (dt/5)*(g_C*C_L*(1-C_L-C_D-M_S-M_C) + gamma_C*(1-C_L-C_D-M_S-M_C) - g_M_C_L*C_L*(M_S + M_C) - d_C*C_L) 
  dC_D <- (dt/5)*(d_C*C_L + M_C*(h_M_C + d_M) - g_M*C_D*(M_S + M_C) - e*C_D)
  dM_S <- (dt/5)*(g_M*M_S*(1-C_L-C_D-M_S-M_C) + g_M*M_C*(1-C_L-C_D-M_S-M_C) + gamma_M*(1-C_L-C_D-M_S-M_C) - M_S*(h_M_S + d_M))
  dM_C <- (dt/5)*(g_M*M_C*C_D + g_M*M_S*C_D + g_M_C_L*C_L*(M_S + M_C) + gamma_M*C_D - M_C*(h_M_C + d_M + e))
  
  C_L.bio[i] <- C_L + dC_L
  C_D.bio[i] <- C_D + dC_D
  M_S.bio[i] <- M_S + dM_S
  M_C.bio[i] <- M_C + dM_C
  
  }
  
}

# Plot of time series for all space holders
plot(tset_bio_dist, C_L.bio, col = "blue", lwd = 1, lty = 'solid', type = 'l', xlab = "Time (months)", ylab = "Proportional cover", xlim = c(0, max(tset_bio_dist)), ylim = c(0, 1))
lines(tset_bio_dist, C_D.bio, col = "black", lwd = 1, lty = 'solid')
lines(tset_bio_dist, M_S.bio, col = "green", lwd = 1, lty = 'solid')
lines(tset_bio_dist, M_C.bio, col = "grey", lwd = 1, lty = 'solid')
#legend(x=350,y=0.7,legend=c('Live coral','Dead coral','Macroalgae (open substrate)','Macroalgae  (dead coral)'), lwd=1, lty='solid', col=c('blue','black','green','grey'))
#abline(v=100, lty = 'dashed')
#abline(v=220, lty = 'dashed')
#abline(v=340, lty = 'dashed')
#abline(v=460, lty = 'dashed')

```

Equilibrium cover following structure removing disturbances, ten year frequency
```{r Equilibrium cover structure-removing ten years, echo=FALSE}


# create holding vectors for equilibration time following disturbance for all space holders
C_L.phys.eq_cover.ten_year <- NaN*disturb_set
C_D.phys.eq_cover.ten_year <- NaN*disturb_set
M_S.phys.eq_cover.ten_year <- NaN*disturb_set
M_C.phys.eq_cover.ten_year <- NaN*disturb_set

# create outer for() loop
# for each value j from 1 to the end of disturb_set (so for each value of disturb_set)
for(j in 1:length(disturb_set)){
	# assign value for 'disturb_set' for this iteration Y to that simulation
  disturbance_severity <- disturb_set[j] 
	
  # create holding vectors for the inner loop
	C_L.phys.ten_year <- NaN*tset.disturb 
	C_L.phys.ten_year[1] <- 0.6
	C_D.phys.ten_year <- NaN*tset.disturb 
	C_D.phys.ten_year[1] <- 0
	M_S.phys.ten_year <- NaN*tset.disturb 
	M_S.phys.ten_year[1] <- 0.05
	M_C.phys.ten_year <- NaN*tset.disturb 
	M_C.phys.ten_year[1] <- 0.05

# Inner loop	
# for each value of 'disturb_set'
  for(i in 2:length(tset.disturb)){
  	
   if(tset.disturb[i] !=t_disturb_1 & tset.disturb[i] !=t_disturb_3 & tset.disturb[i] !=t_disturb_5 & tset.disturb[i] !=t_disturb_7 & tset.disturb[i] !=t_disturb_9 & tset.disturb[i] !=t_disturb_11 & tset.disturb[i] !=t_disturb_13 & tset.disturb[i] !=t_disturb_15 & tset.disturb[i] !=t_disturb_17){ # change in time
    
    # use variable assignment to create objects to plug into calculations 
  	C_L <- C_L.phys.ten_year[i-1]
  	C_D <- C_D.phys.ten_year[i-1]
    M_S <- M_S.phys.ten_year[i-1]
  	M_C <- M_C.phys.ten_year[i-1]
  	
  	# calculate change in proportional cover at each timestep
    dC_L <- (dt/5)*(g_C*C_L*(1-C_L-C_D-M_S-M_C) + gamma_C*(1-C_L-C_D-M_S-M_C) - g_M_C_L*C_L*(M_S + M_C) - d_C*C_L) 
    dC_D <- (dt/5)*(d_C*C_L + M_C*(h_M_C + d_M) - g_M*C_D*(M_S + M_C) - e*C_D)
    dM_S <- (dt/5)*(g_M*M_S*(1-C_L-C_D-M_S-M_C) + g_M*M_C*(1-C_L-C_D-M_S-M_C) + gamma_M*(1-C_L-C_D-M_S-M_C) - M_S*(h_M_S + d_M))
    dM_C <- (dt/5)*(g_M*M_C*C_D + g_M*M_S*C_D + g_M_C_L*C_L*(M_S + M_C) + gamma_M*C_D - M_C*(h_M_C + d_M + e))
  	
  	
  	# calculate the population density at each timestep
  	C_L.phys.ten_year[i] <- C_L + dC_L
  	C_D.phys.ten_year[i] <- C_D + dC_D
  	M_S.phys.ten_year[i] <- M_S + dM_S
  	M_C.phys.ten_year[i] <- M_C + dM_C
    }
    
     else {
        dt <- tset.disturb[i] - tset.disturb[i-1] # change in time
    
    # use variable assignment to create objects to plug into calculations 
  	C_L <- C_L.phys.ten_year[i-1]*(1-disturbance_severity)
  	C_D <- C_D.phys.ten_year[i-1]*(1-disturbance_severity)
    M_S <- M_S.phys.ten_year[i-1]*(1-disturbance_severity)
  	M_C <- M_C.phys.ten_year[i-1]*(1-disturbance_severity)
  	
  	# calculate change in proportional cover at each timestep
    dC_L <- (dt/5)*(g_C*C_L*(1-C_L-C_D-M_S-M_C) + gamma_C*(1-C_L-C_D-M_S-M_C) - g_M_C_L*C_L*(M_S + M_C) - d_C*C_L) 
    dC_D <- (dt/5)*(d_C*C_L + M_C*(h_M_C + d_M) - g_M*C_D*(M_S + M_C) - e*C_D)
    dM_S <- (dt/5)*(g_M*M_S*(1-C_L-C_D-M_S-M_C) + g_M*M_C*(1-C_L-C_D-M_S-M_C) + gamma_M*(1-C_L-C_D-M_S-M_C) - M_S*(h_M_S + d_M))
    dM_C <- (dt/5)*(g_M*M_C*C_D + g_M*M_S*C_D + g_M_C_L*C_L*(M_S + M_C) + gamma_M*C_D - M_C*(h_M_C + d_M + e))
  	
  	
  	# calculate the proportional cover at each timestep
  	C_L.phys.ten_year[i] <- C_L + dC_L
  	C_D.phys.ten_year[i] <- C_D + dC_D
  	M_S.phys.ten_year[i] <- M_S + dM_S
  	M_C.phys.ten_year[i] <- M_C + dM_C
    }
  }
# final output is equilibrium cover for a given severity of disturbance
C_L.phys.eq_cover.ten_year[j] <- C_L.phys.ten_year[max(tset.disturb)]
C_D.phys.eq_cover.ten_year[j] <- C_D.phys.ten_year[max(tset.disturb)]
M_S.phys.eq_cover.ten_year[j] <- M_S.phys.ten_year[max(tset.disturb)]
M_C.phys.eq_cover.ten_year[j] <- M_C.phys.ten_year[max(tset.disturb)]
}

```

```{r Equilibrium cover structure-retaining ten years, echo=FALSE}

# create holding vectors for equilibration time following disturbance for all space holders
C_L.bio.eq_cover.ten_year <- NaN*disturb_set
C_D.bio.eq_cover.ten_year <- NaN*disturb_set
M_S.bio.eq_cover.ten_year <- NaN*disturb_set
M_C.bio.eq_cover.ten_year <- NaN*disturb_set

# create outer for() loop
# for each value j from 1 to the end of disturb_set (so for each value of disturb_set)
for(j in 1:length(disturb_set)){
	# assign value for 'disturb_set' for this iteration Y to that simulation
  disturbance_severity <- disturb_set[j] 
	
  # create holding vectors for the inner loop
	C_L.bio.ten_year <- NaN*tset.disturb 
	C_L.bio.ten_year[1] <- 0.6
	C_D.bio.ten_year <- NaN*tset.disturb 
	C_D.bio.ten_year[1] <- 0
	M_S.bio.ten_year <- NaN*tset.disturb 
	M_S.bio.ten_year[1] <- 0.05
	M_C.bio.ten_year <- NaN*tset.disturb 
	M_C.bio.ten_year[1] <- 0.05

# Inner loop	
# for each value of 'disturb_set'
	for(i in 2:length(tset.disturb)){
	  
  if(tset.disturb[i] !=t_disturb_1 & tset.disturb[i] !=t_disturb_3 & tset.disturb[i] !=t_disturb_5 & tset.disturb[i] !=t_disturb_7 & tset.disturb[i] !=t_disturb_9 & tset.disturb[i] !=t_disturb_11 & tset.disturb[i] !=t_disturb_13 & tset.disturb[i] !=t_disturb_15 & tset.disturb[i] !=t_disturb_17){ # change in time
    
    # use variable assignment to create objects to plug into calculations 
  	C_L <- C_L.bio.ten_year[i-1]
  	C_D <- C_D.bio.ten_year[i-1]
    M_S <- M_S.bio.ten_year[i-1]
  	M_C <- M_C.bio.ten_year[i-1]
  	
  	# calculate change in proportional cover at each timestep
    dC_L <- (dt/5)*(g_C*C_L*(1-C_L-C_D-M_S-M_C) + gamma_C*(1-C_L-C_D-M_S-M_C) - g_M_C_L*C_L*(M_S + M_C) - d_C*C_L) 
    dC_D <- (dt/5)*(d_C*C_L + M_C*(h_M_C + d_M) - g_M*C_D*(M_S + M_C) - e*C_D)
    dM_S <- (dt/5)*(g_M*M_S*(1-C_L-C_D-M_S-M_C) + g_M*M_C*(1-C_L-C_D-M_S-M_C) + gamma_M*(1-C_L-C_D-M_S-M_C) - M_S*(h_M_S + d_M))
    dM_C <- (dt/5)*(g_M*M_C*C_D + g_M*M_S*C_D + g_M_C_L*C_L*(M_S + M_C) + gamma_M*C_D - M_C*(h_M_C + d_M + e))
  	
  	
  	# calculate the population density at each timestep
  	C_L.bio.ten_year[i] <- C_L + dC_L
  	C_D.bio.ten_year[i] <- C_D + dC_D
  	M_S.bio.ten_year[i] <- M_S + dM_S
  	M_C.bio.ten_year[i] <- M_C + dM_C
    }
    
     else {
        dt <- tset.disturb[i] - tset.disturb[i-1] # change in time
    
    # use variable assignment to create objects to plug into calculations 
  	C_L <- C_L.bio.ten_year[i-1]*(1-disturbance_severity)
  	C_D <- C_D.bio.ten_year[i-1] + (C_L.bio.ten_year[i-1]-C_L)
    M_S <- M_S.bio.ten_year[i-1]
  	M_C <- M_C.bio.ten_year[i-1]
  	
  	# calculate change in proportional cover at each timestep
    dC_L <- (dt/5)*(g_C*C_L*(1-C_L-C_D-M_S-M_C) + gamma_C*(1-C_L-C_D-M_S-M_C) - g_M_C_L*C_L*(M_S + M_C) - d_C*C_L) 
    dC_D <- (dt/5)*(d_C*C_L + M_C*(h_M_C + d_M) - g_M*C_D*(M_S + M_C) - e*C_D)
    dM_S <- (dt/5)*(g_M*M_S*(1-C_L-C_D-M_S-M_C) + g_M*M_C*(1-C_L-C_D-M_S-M_C) + gamma_M*(1-C_L-C_D-M_S-M_C) - M_S*(h_M_S + d_M))
    dM_C <- (dt/5)*(g_M*M_C*C_D + g_M*M_S*C_D + g_M_C_L*C_L*(M_S + M_C) + gamma_M*C_D - M_C*(h_M_C + d_M + e))
  	
  	
  	# calculate the proportional cover at each timestep
  	C_L.bio.ten_year[i] <- C_L + dC_L
  	C_D.bio.ten_year[i] <- C_D + dC_D
  	M_S.bio.ten_year[i] <- M_S + dM_S
  	M_C.bio.ten_year[i] <- M_C + dM_C
     }
  }
# final output is maximum cover for a given severity of disturbance
C_L.bio.eq_cover.ten_year[j] <- C_L.bio.ten_year[max(tset.disturb)]
C_D.bio.eq_cover.ten_year[j] <- C_D.bio.ten_year[max(tset.disturb)]
M_S.bio.eq_cover.ten_year[j] <- M_S.bio.ten_year[max(tset.disturb)]
M_C.bio.eq_cover.ten_year[j] <- M_C.bio.ten_year[max(tset.disturb)]
}

```

Equilibrium cover five year frequency
```{r Equilibrium cover structure-removing five years, echo=FALSE}


# create holding vectors for equilibration time following disturbance for all space holders
C_L.phys.eq_cover.five_year <- NaN*disturb_set
C_D.phys.eq_cover.five_year <- NaN*disturb_set
M_S.phys.eq_cover.five_year <- NaN*disturb_set
M_C.phys.eq_cover.five_year <- NaN*disturb_set

# create outer for() loop
# for each value j from 1 to the end of disturb_set (so for each value of disturb_set)
for(j in 1:length(disturb_set)){
	# assign value for 'disturb_set' for this iteration Y to that simulation
  disturbance_severity <- disturb_set[j] 
	
  # create holding vectors for the inner loop
	C_L.phys.five_year <- NaN*tset.disturb 
	C_L.phys.five_year[1] <- 0.6
	C_D.phys.five_year <- NaN*tset.disturb 
	C_D.phys.five_year[1] <- 0
	M_S.phys.five_year <- NaN*tset.disturb 
	M_S.phys.five_year[1] <- 0.05
	M_C.phys.five_year <- NaN*tset.disturb 
	M_C.phys.five_year[1] <- 0.05

# Inner loop	
# for each value of 'disturb_set'
  for(i in 2:length(tset.disturb)){
  	
   if(tset.disturb[i] !=t_disturb_1 & tset.disturb[i] !=t_disturb_3 & tset.disturb[i] !=t_disturb_5 & tset.disturb[i] !=t_disturb_7 & tset.disturb[i] !=t_disturb_9 & tset.disturb[i] !=t_disturb_11 & tset.disturb[i] !=t_disturb_13 & tset.disturb[i] !=t_disturb_15 & tset.disturb[i] !=t_disturb_17 & tset.disturb[i] !=t_disturb_2 & tset.disturb[i] !=t_disturb_4 & tset.disturb[i] !=t_disturb_6 & tset.disturb[i] !=t_disturb_8 & tset.disturb[i] !=t_disturb_10 & tset.disturb[i] !=t_disturb_12 & tset.disturb[i] !=t_disturb_14 & tset.disturb[i] !=t_disturb_16 & tset.disturb[i] !=t_disturb_18){ # change in time
    
    # use variable assignment to create objects to plug into calculations 
  	C_L <- C_L.phys.five_year[i-1]
  	C_D <- C_D.phys.five_year[i-1]
    M_S <- M_S.phys.five_year[i-1]
  	M_C <- M_C.phys.five_year[i-1]
  	
  	# calculate change in proportional cover at each timestep
    dC_L <- (dt/5)*(g_C*C_L*(1-C_L-C_D-M_S-M_C) + gamma_C*(1-C_L-C_D-M_S-M_C) - g_M_C_L*C_L*(M_S + M_C) - d_C*C_L) 
    dC_D <- (dt/5)*(d_C*C_L + M_C*(h_M_C + d_M) - g_M*C_D*(M_S + M_C) - e*C_D)
    dM_S <- (dt/5)*(g_M*M_S*(1-C_L-C_D-M_S-M_C) + g_M*M_C*(1-C_L-C_D-M_S-M_C) + gamma_M*(1-C_L-C_D-M_S-M_C) - M_S*(h_M_S + d_M))
    dM_C <- (dt/5)*(g_M*M_C*C_D + g_M*M_S*C_D + g_M_C_L*C_L*(M_S + M_C) + gamma_M*C_D - M_C*(h_M_C + d_M + e))
  	
  	
  	# calculate the population density at each timestep
  	C_L.phys.five_year[i] <- C_L + dC_L
  	C_D.phys.five_year[i] <- C_D + dC_D
  	M_S.phys.five_year[i] <- M_S + dM_S
  	M_C.phys.five_year[i] <- M_C + dM_C
    }
    
     else {
        dt <- tset.disturb[i] - tset.disturb[i-1] # change in time
    
    # use variable assignment to create objects to plug into calculations 
  	C_L <- C_L.phys.five_year[i-1]*(1-disturbance_severity)
  	C_D <- C_D.phys.five_year[i-1]*(1-disturbance_severity)
    M_S <- M_S.phys.five_year[i-1]*(1-disturbance_severity)
  	M_C <- M_C.phys.five_year[i-1]*(1-disturbance_severity)
  	
  	# calculate change in proportional cover at each timestep
    dC_L <- (dt/5)*(g_C*C_L*(1-C_L-C_D-M_S-M_C) + gamma_C*(1-C_L-C_D-M_S-M_C) - g_M_C_L*C_L*(M_S + M_C) - d_C*C_L) 
    dC_D <- (dt/5)*(d_C*C_L + M_C*(h_M_C + d_M) - g_M*C_D*(M_S + M_C) - e*C_D)
    dM_S <- (dt/5)*(g_M*M_S*(1-C_L-C_D-M_S-M_C) + g_M*M_C*(1-C_L-C_D-M_S-M_C) + gamma_M*(1-C_L-C_D-M_S-M_C) - M_S*(h_M_S + d_M))
    dM_C <- (dt/5)*(g_M*M_C*C_D + g_M*M_S*C_D + g_M_C_L*C_L*(M_S + M_C) + gamma_M*C_D - M_C*(h_M_C + d_M + e))
  	
  	
  	# calculate the proportional cover at each timestep
  	C_L.phys.five_year[i] <- C_L + dC_L
  	C_D.phys.five_year[i] <- C_D + dC_D
  	M_S.phys.five_year[i] <- M_S + dM_S
  	M_C.phys.five_year[i] <- M_C + dM_C
    }
  }
# final output is equilibrium cover for a given severity of disturbance
C_L.phys.eq_cover.five_year[j] <- C_L.phys.five_year[max(tset.disturb)]
C_D.phys.eq_cover.five_year[j] <- C_D.phys.five_year[max(tset.disturb)]
M_S.phys.eq_cover.five_year[j] <- M_S.phys.five_year[max(tset.disturb)]
M_C.phys.eq_cover.five_year[j] <- M_C.phys.five_year[max(tset.disturb)]
}

```

```{r Equilibrium cover structure-retaining five years, echo=FALSE}

# create holding vectors for equilibration time following disturbance for all space holders
C_L.bio.eq_cover.five_year <- NaN*disturb_set
C_D.bio.eq_cover.five_year <- NaN*disturb_set
M_S.bio.eq_cover.five_year <- NaN*disturb_set
M_C.bio.eq_cover.five_year <- NaN*disturb_set

# create outer for() loop
# for each value j from 1 to the end of disturb_set (so for each value of disturb_set)
for(j in 1:length(disturb_set)){
	# assign value for 'disturb_set' for this iteration Y to that simulation
  disturbance_severity <- disturb_set[j] 
	
  # create holding vectors for the inner loop
	C_L.bio.five_year <- NaN*tset.disturb 
	C_L.bio.five_year[1] <- 0.6
	C_D.bio.five_year <- NaN*tset.disturb 
	C_D.bio.five_year[1] <- 0
	M_S.bio.five_year <- NaN*tset.disturb 
	M_S.bio.five_year[1] <- 0.05
	M_C.bio.five_year <- NaN*tset.disturb 
	M_C.bio.five_year[1] <- 0.05

# Inner loop	
# for each value of 'disturb_set'
	for(i in 2:length(tset.disturb)){
	  
  if(tset.disturb[i] !=t_disturb_1 & tset.disturb[i] !=t_disturb_3 & tset.disturb[i] !=t_disturb_5 & tset.disturb[i] !=t_disturb_7 & tset.disturb[i] !=t_disturb_9 & tset.disturb[i] !=t_disturb_11 & tset.disturb[i] !=t_disturb_13 & tset.disturb[i] !=t_disturb_15 & tset.disturb[i] !=t_disturb_17 & tset.disturb[i] !=t_disturb_2 & tset.disturb[i] !=t_disturb_4 & tset.disturb[i] !=t_disturb_6 & tset.disturb[i] !=t_disturb_8 & tset.disturb[i] !=t_disturb_10 & tset.disturb[i] !=t_disturb_12 & tset.disturb[i] !=t_disturb_14 & tset.disturb[i] !=t_disturb_16 & tset.disturb[i] !=t_disturb_18){ # change in time
    
    # use variable assignment to create objects to plug into calculations 
  	C_L <- C_L.bio.five_year[i-1]
  	C_D <- C_D.bio.five_year[i-1]
    M_S <- M_S.bio.five_year[i-1]
  	M_C <- M_C.bio.five_year[i-1]
  	
  	# calculate change in proportional cover at each timestep
    dC_L <- (dt/5)*(g_C*C_L*(1-C_L-C_D-M_S-M_C) + gamma_C*(1-C_L-C_D-M_S-M_C) - g_M_C_L*C_L*(M_S + M_C) - d_C*C_L) 
    dC_D <- (dt/5)*(d_C*C_L + M_C*(h_M_C + d_M) - g_M*C_D*(M_S + M_C) - e*C_D)
    dM_S <- (dt/5)*(g_M*M_S*(1-C_L-C_D-M_S-M_C) + g_M*M_C*(1-C_L-C_D-M_S-M_C) + gamma_M*(1-C_L-C_D-M_S-M_C) - M_S*(h_M_S + d_M))
    dM_C <- (dt/5)*(g_M*M_C*C_D + g_M*M_S*C_D + g_M_C_L*C_L*(M_S + M_C) + gamma_M*C_D - M_C*(h_M_C + d_M + e))
  	
  	
  	# calculate the population density at each timestep
  	C_L.bio.five_year[i] <- C_L + dC_L
  	C_D.bio.five_year[i] <- C_D + dC_D
  	M_S.bio.five_year[i] <- M_S + dM_S
  	M_C.bio.five_year[i] <- M_C + dM_C
    }
    
     else {
        dt <- tset.disturb[i] - tset.disturb[i-1] # change in time
    
    # use variable assignment to create objects to plug into calculations 
  	C_L <- C_L.bio.five_year[i-1]*(1-disturbance_severity)
  	C_D <- C_D.bio.five_year[i-1] + (C_L.bio.five_year[i-1]-C_L)
    M_S <- M_S.bio.five_year[i-1]
  	M_C <- M_C.bio.five_year[i-1]
  	
  	# calculate change in proportional cover at each timestep
    dC_L <- (dt/5)*(g_C*C_L*(1-C_L-C_D-M_S-M_C) + gamma_C*(1-C_L-C_D-M_S-M_C) - g_M_C_L*C_L*(M_S + M_C) - d_C*C_L) 
    dC_D <- (dt/5)*(d_C*C_L + M_C*(h_M_C + d_M) - g_M*C_D*(M_S + M_C) - e*C_D)
    dM_S <- (dt/5)*(g_M*M_S*(1-C_L-C_D-M_S-M_C) + g_M*M_C*(1-C_L-C_D-M_S-M_C) + gamma_M*(1-C_L-C_D-M_S-M_C) - M_S*(h_M_S + d_M))
    dM_C <- (dt/5)*(g_M*M_C*C_D + g_M*M_S*C_D + g_M_C_L*C_L*(M_S + M_C) + gamma_M*C_D - M_C*(h_M_C + d_M + e))
  	
  	
  	# calculate the proportional cover at each timestep
  	C_L.bio.five_year[i] <- C_L + dC_L
  	C_D.bio.five_year[i] <- C_D + dC_D
  	M_S.bio.five_year[i] <- M_S + dM_S
  	M_C.bio.five_year[i] <- M_C + dM_C
     }
  }
# final output is maximum cover for a given severity of disturbance
C_L.bio.eq_cover.five_year[j] <- C_L.bio.five_year[max(tset.disturb)]
C_D.bio.eq_cover.five_year[j] <- C_D.bio.five_year[max(tset.disturb)]
M_S.bio.eq_cover.five_year[j] <- M_S.bio.five_year[max(tset.disturb)]
M_C.bio.eq_cover.five_year[j] <- M_C.bio.five_year[max(tset.disturb)]
}

```

Equilibrium cover plots
```{r Equilibrium cover plots, echo=FALSE}

# Live coral
plot(disturb_set, C_L.phys.eq_cover.ten_year, col = "blue", lwd = 2, lty = 'solid', type = 'l', xlab = "Severity of disturbance", ylab = "Equilibrium cover", xlim = c(0, 1), ylim = c(0, 1.1), main = "Live coral")
lines(disturb_set, C_L.bio.eq_cover.ten_year, col = "red", lwd = 2, lty = 'solid', type = 'l')
lines(disturb_set, C_L.phys.eq_cover.five_year, col = "blue", lwd = 2, lty = 'dotted', type = 'l')
lines(disturb_set, C_L.bio.eq_cover.five_year, col = "red", lwd = 2, lty = 'dotted', type = 'l')
#legend(x=0.6,y=1.1,legend=c('Structure-removing disturbance','Structure-retaining disturbance','10-year frequency','5-year frequency'),lwd=1,lty=c('solid','solid','dotted','dotted'),col=c('blue','red', 'black','black'))

# Total macroalgae
plot(disturb_set, c(M_C.phys.eq_cover.ten_year + M_S.phys.eq_cover.ten_year), col = "blue", lwd = 2, lty = 'solid', type = 'l', xlab = "Severity of disturbance", ylab = "Equilibrium cover", xlim = c(0, 1), ylim = c(0, 1.1), main = "Total macroalgae")
lines(disturb_set, c(M_C.bio.eq_cover.ten_year + M_S.bio.eq_cover.ten_year), col = "red", lwd = 2, lty = 'solid', type = 'l')
lines(disturb_set, c(M_C.phys.eq_cover.five_year + M_S.phys.eq_cover.five_year), col = "blue", lwd = 2, lty = 'dotted', type = 'l')
lines(disturb_set, c(M_C.bio.eq_cover.five_year + M_S.bio.eq_cover.five_year), col = "red", lwd = 2, lty = 'dotted', type = 'l')
#legend(x=0.6,y=1.1,legend=c('Structure-removing','Structure-retaining'), lwd=1, lty='solid', col=c('blue','red'))

```

Structure-removing disturbances: 50% removal of all space holders every five years
```{r 50% structure-removing disturbance five year frequency, echo = FALSE}

# Create time vector, holding vectors for each state variable, and set initial conditions
tset_phys_dist <- seq(from = 0, to = 1200)
C_L.simu_phys <- NaN*tset_phys_dist; C_L.simu_phys[1] <- 0.6
C_D.simu_phys <- NaN*tset_phys_dist; C_D.simu_phys[1] <- 0
M_S.simu_phys <- NaN*tset_phys_dist; M_S.simu_phys[1] <- 0.05
M_C.simu_phys <- NaN*tset_phys_dist; M_C.simu_phys[1] <- 0.05

# For() loop with if/else statement to simulate trajectories of spatial occupancy by live coral and macroalgae before and after disturbance

for(i in 2:length(tset_phys_dist)){ # For every element after the first in the time vector...
  
  if(tset_phys_dist[i] !=t_disturb_10 & tset_phys_dist[i] !=t_disturb_11 & tset_phys_dist[i] !=t_disturb_12 & tset_phys_dist[i] !=t_disturb_13 & tset_phys_dist[i] !=t_disturb_14 & tset_phys_dist[i] !=t_disturb_15 & tset_phys_dist[i] !=t_disturb_16 & tset_phys_dist[i] !=t_disturb_17 & tset_phys_dist[i] !=t_disturb_18 & tset_phys_dist[i] !=t_disturb_19 & tset_phys_dist[i] !=t_disturb_20 & tset_phys_dist[i] !=t_disturb_21 & tset_phys_dist[i] !=t_disturb_22 & tset_phys_dist[i] !=t_disturb_23 & tset_phys_dist[i] !=t_disturb_24 & tset_phys_dist[i] !=t_disturb_25 & tset_phys_dist[i] !=t_disturb_26 & tset_phys_dist[i] !=t_disturb_27){ # ...if the time point does not equal the time of disturbance, run the simulation as normal...
  
  dt <- tset_phys_dist[i] - tset_phys_dist[i-1] # calculate change in time  
 
  # Proportional cover at previous time step  
  C_L <- C_L.simu_phys[i-1]
  C_D <- C_D.simu_phys[i-1]
  M_S <- M_S.simu_phys[i-1]
  M_C <- M_C.simu_phys[i-1]
  
  # Change in proportional cover, converted to units of months (/5)
  dC_L <- (dt/5)*(g_C*C_L*(1-C_L-C_D-M_S-M_C) + gamma_C*(1-C_L-C_D-M_S-M_C) - g_M_C_L*C_L*(M_S + M_C) - d_C*C_L) 
  dC_D <- (dt/5)*(d_C*C_L + M_C*(h_M_C + d_M) - g_M*C_D*(M_S + M_C) - e*C_D)
  dM_S <- (dt/5)*(g_M*M_S*(1-C_L-C_D-M_S-M_C) + g_M*M_C*(1-C_L-C_D-M_S-M_C) + gamma_M*(1-C_L-C_D-M_S-M_C) - M_S*(h_M_S + d_M))
  dM_C <- (dt/5)*(g_M*M_C*C_D + g_M*M_S*C_D + g_M_C_L*C_L*(M_S + M_C) + gamma_M*C_D - M_C*(h_M_C + d_M + e))
 
  # Proportional cover at new time step   
  C_L.simu_phys[i] <- C_L.simu_phys[i-1] + dC_L
  C_D.simu_phys[i] <- C_D.simu_phys[i-1] + dC_D
  M_S.simu_phys[i] <- M_S.simu_phys[i-1] + dM_S
  M_C.simu_phys[i] <- M_C.simu_phys[i-1] + dM_C
  }
  
    else { # ... but if the time step does equal the time of disturbance...
  
  dt <- tset_phys_dist[i] - tset_phys_dist[i-1]

  # reduce all space holders to 50% of their pre-disturbance levels
  C_L <- C_L.simu_phys[i-1]*0.5
  C_D <- C_D.simu_phys[i-1]*0.5
  M_S <- M_S.simu_phys[i-1]*0.5
  M_C <- M_C.simu_phys[i-1]*0.5
  
  dC_L <- (dt/5)*(g_C*C_L*(1-C_L-C_D-M_S-M_C) + gamma_C*(1-C_L-C_D-M_S-M_C) - g_M_C_L*C_L*(M_S + M_C) - d_C*C_L) 
  dC_D <- (dt/5)*(d_C*C_L + M_C*(h_M_C + d_M) - g_M*C_D*(M_S + M_C) - e*C_D)
  dM_S <- (dt/5)*(g_M*M_S*(1-C_L-C_D-M_S-M_C) + g_M*M_C*(1-C_L-C_D-M_S-M_C) + gamma_M*(1-C_L-C_D-M_S-M_C) - M_S*(h_M_S + d_M))
  dM_C <- (dt/5)*(g_M*M_C*C_D + g_M*M_S*C_D + g_M_C_L*C_L*(M_S + M_C) + gamma_M*C_D - M_C*(h_M_C + d_M + e))
  
  C_L.simu_phys[i] <- C_L+ dC_L
  C_D.simu_phys[i] <- C_D+ dC_D
  M_S.simu_phys[i] <- M_S+ dM_S
  M_C.simu_phys[i] <- M_C+ dM_C
  }
  
}

# Plot of time series
plot(tset_phys_dist, C_L.simu_phys, col = "blue", lwd = 1, lty = 'solid', type = 'l', xlab = "Time (months)", ylab = "Proportional cover", xlim = c(0, max(tset_phys_dist)), ylim = c(0, 1))
lines(tset_phys_dist, C_D.simu_phys, col = "black", lwd = 1, lty = 'solid')
lines(tset_phys_dist, M_S.simu_phys, col = "green", lwd = 1, lty = 'solid')
lines(tset_phys_dist, M_C.simu_phys, col = "grey", lwd = 1, lty = 'solid')
#legend(x=350,y=0.7,legend=c('Live coral','Dead coral','Macroalgae (open substrate)','Macroalgae (dead coral)'),lwd=1,lty='solid',col=c('blue','black','green','grey'))
#abline(v=100, lty = 'dashed')
#abline(v=340, lty = 'dashed')
#abline(v=580, lty = 'dashed')
#abline(v=820, lty = 'dashed')
```

50% reduction in live coral cover that is transferred to dead coral every five years; macroalgae unaffected
```{r 50% structure-retaining disturbance five year frequency, echo = FALSE}

# Create time vector, holding vectors for each state variable, and set initial conditions
tset_bio_dist <- seq(from = 0, to = 1200)
C_L.bio <- NaN*tset_bio_dist; C_L.bio[1] <- 0.6
C_D.bio <- NaN*tset_bio_dist; C_D.bio[1] <- 0
M_S.bio <- NaN*tset_bio_dist; M_S.bio[1] <- 0.05
M_C.bio <- NaN*tset_bio_dist; M_C.bio[1] <- 0.05

# For() loop with if/else statement to simulate trajectories of spatial occupancy by live coral and macroalgae before and after disturbance
for(i in 2:length(tset_bio_dist)){ # For every element after the first in the time vector...
  
 if(tset_bio_dist[i] !=t_disturb_10 & tset_bio_dist[i] !=t_disturb_11 & tset_bio_dist[i] !=t_disturb_12 & tset_bio_dist[i] !=t_disturb_13 & tset_bio_dist[i] !=t_disturb_14 & tset_bio_dist[i] !=t_disturb_15 & tset_bio_dist[i] !=t_disturb_16 & tset_bio_dist[i] !=t_disturb_17 & tset_bio_dist[i] !=t_disturb_18 & tset_bio_dist[i] !=t_disturb_19 & tset_bio_dist[i] !=t_disturb_20 & tset_bio_dist[i] !=t_disturb_21 & tset_bio_dist[i] !=t_disturb_22 & tset_bio_dist[i] !=t_disturb_23 & tset_bio_dist[i] !=t_disturb_24 & tset_bio_dist[i] !=t_disturb_25 & tset_bio_dist[i] !=t_disturb_26 & tset_bio_dist[i] !=t_disturb_27){ # ...if the time point does not equal the time of disturbance, run the simulation as normal...

# calculate change in time  
  dt <- tset_bio_dist[i] - tset_bio_dist[i-1] 

  # Proportional cover at previous time step  
  C_L <- C_L.bio[i-1] 
  C_D <- C_D.bio[i-1]
  M_S <- M_S.bio[i-1]
  M_C <- M_C.bio[i-1]
 
# Change in proportional cover, converted to units of months (/5)
  dC_L <- (dt/5)*(g_C*C_L*(1-C_L-C_D-M_S-M_C) + gamma_C*(1-C_L-C_D-M_S-M_C) - g_M_C_L*C_L*(M_S + M_C) - d_C*C_L) 
  dC_D <- (dt/5)*(d_C*C_L + M_C*(h_M_C + d_M) - g_M*C_D*(M_S + M_C) - e*C_D)
  dM_S <- (dt/5)*(g_M*M_S*(1-C_L-C_D-M_S-M_C) + g_M*M_C*(1-C_L-C_D-M_S-M_C) + gamma_M*(1-C_L-C_D-M_S-M_C) - M_S*(h_M_S + d_M))
  dM_C <- (dt/5)*(g_M*M_C*C_D + g_M*M_S*C_D + g_M_C_L*C_L*(M_S + M_C) + gamma_M*C_D - M_C*(h_M_C + d_M + e))

# Proportional cover at new time step  
  C_L.bio[i] <- C_L.bio[i-1] + dC_L
  C_D.bio[i] <- C_D.bio[i-1] + dC_D
  M_S.bio[i] <- M_S.bio[i-1] + dM_S
  M_C.bio[i] <- M_C.bio[i-1] + dM_C
  
  }

  # ... but if the time step does equal the time of disturbance...
    else {
  
  dt <- tset_bio_dist[i] - tset_bio_dist[i-1]

  C_L <- C_L.bio[i-1]*0.5 # reduce live coral to 50% of its pre-disturbance level, and transfer this directly to the dead coral class
  C_D <- C_D.bio[i-1] + (C_L.bio[i-1]-C_L)
  M_S <- M_S.bio[i-1]
  M_C <- M_C.bio[i-1]
  
  dC_L <- (dt/5)*(g_C*C_L*(1-C_L-C_D-M_S-M_C) + gamma_C*(1-C_L-C_D-M_S-M_C) - g_M_C_L*C_L*(M_S + M_C) - d_C*C_L) 
  dC_D <- (dt/5)*(d_C*C_L + M_C*(h_M_C + d_M) - g_M*C_D*(M_S + M_C) - e*C_D)
  dM_S <- (dt/5)*(g_M*M_S*(1-C_L-C_D-M_S-M_C) + g_M*M_C*(1-C_L-C_D-M_S-M_C) + gamma_M*(1-C_L-C_D-M_S-M_C) - M_S*(h_M_S + d_M))
  dM_C <- (dt/5)*(g_M*M_C*C_D + g_M*M_S*C_D + g_M_C_L*C_L*(M_S + M_C) + gamma_M*C_D - M_C*(h_M_C + d_M + e))
  
  C_L.bio[i] <- C_L + dC_L
  C_D.bio[i] <- C_D + dC_D
  M_S.bio[i] <- M_S + dM_S
  M_C.bio[i] <- M_C + dM_C
  
  }
  
}

# Plot of time series for all space holders
plot(tset_bio_dist, C_L.bio, col = "blue", lwd = 1, lty = 'solid', type = 'l', xlab = "Time (months)", ylab = "Proportional cover", xlim = c(0, max(tset_bio_dist)), ylim = c(0, 1))
lines(tset_bio_dist, C_D.bio, col = "black", lwd = 1, lty = 'solid')
lines(tset_bio_dist, M_S.bio, col = "green", lwd = 1, lty = 'solid')
lines(tset_bio_dist, M_C.bio, col = "grey", lwd = 1, lty = 'solid')
#legend(x=350,y=0.7,legend=c('Live coral','Dead coral','Macroalgae (open substrate)','Macroalgae  (dead coral)'), lwd=1, lty='solid', col=c('blue','black','green','grey'))
#abline(v=100, lty = 'dashed')
#abline(v=340, lty = 'dashed')
#abline(v=580, lty = 'dashed')
#abline(v=820, lty = 'dashed')

```

#### Equilibrium cover after 100 years of live coral and total macroalgae: ten year frequency

Equilibrium cover after 100 years
```{r Equilibrium cover structure-retaining ten years, echo=FALSE}

disturb_set <- seq(from = 0, to = 1, by = .01) # Time vector for outer loop
tset.disturb <- seq(from = 0, to = 1200) # Time vector for inner loop

# create holding vectors for equilibration time following disturbance for all space holders
C_L.bio.eq_cover <- NaN*disturb_set
C_D.bio.eq_cover <- NaN*disturb_set
M_S.bio.eq_cover <- NaN*disturb_set
M_C.bio.eq_cover <- NaN*disturb_set

# create outer for() loop
# for each value j from 1 to the end of disturb_set (so for each value of disturb_set)
for(j in 1:length(disturb_set)){
	# assign value for 'disturb_set' for this iteration Y to that simulation
  disturbance_severity <- disturb_set[j] 
	
  # create holding vectors for the inner loop
	C_L.bio <- NaN*tset.disturb 
	C_L.bio[1] <- 0.6
	C_D.bio <- NaN*tset.disturb 
	C_D.bio[1] <- 0
	M_S.bio <- NaN*tset.disturb 
	M_S.bio[1] <- 0.05
	M_C.bio <- NaN*tset.disturb 
	M_C.bio[1] <- 0.05

# Inner loop	
# for each value of 'disturb_set'
  for(i in 2:length(tset.disturb)){
  	
   if(tset_phys_dist_sev[i] !=t_disturb_1 & tset_phys_dist_sev[i] !=t_disturb_2 & tset_phys_dist_sev[i] !=t_disturb_3 & tset_phys_dist_sev[i] !=t_disturb_4 & tset_phys_dist_sev[i] !=t_disturb_5 & tset_phys_dist_sev[i] !=t_disturb_6 & tset_phys_dist_sev[i] !=t_disturb_7 & tset_phys_dist_sev[i] !=t_disturb_8 & tset_phys_dist_sev[i] !=t_disturb_9){ # change in time
    
    # use variable assignment to create objects to plug into calculations 
  	C_L <- C_L.bio[i-1]
  	C_D <- C_D.bio[i-1]
    M_S <- M_S.bio[i-1]
  	M_C <- M_C.bio[i-1]
  	
  	# calculate change in proportional cover at each timestep
    dC_L <- (dt/5)*(g_C*C_L*(1-C_L-C_D-M_S-M_C) + gamma_C*(1-C_L-C_D-M_S-M_C) - g_M_C_L*C_L*(M_S + M_C) - d_C*C_L) 
    dC_D <- (dt/5)*(d_C*C_L + M_C*(h_M_C + d_M) - g_M*C_D*(M_S + M_C) - e*C_D)
    dM_S <- (dt/5)*(g_M*M_S*(1-C_L-C_D-M_S-M_C) + g_M*M_C*(1-C_L-C_D-M_S-M_C) + gamma_M*(1-C_L-C_D-M_S-M_C) - M_S*(h_M_S + d_M))
    dM_C <- (dt/5)*(g_M*M_C*C_D + g_M*M_S*C_D + g_M_C_L*C_L*(M_S + M_C) + gamma_M*C_D - M_C*(h_M_C + d_M + e))
  	
  	
  	# calculate the population density at each timestep
  	C_L.bio[i] <- C_L + dC_L
  	C_D.bio[i] <- C_D + dC_D
  	M_S.bio[i] <- M_S + dM_S
  	M_C.bio[i] <- M_C + dM_C
    }
    
     else {
        dt <- tset.disturb[i] - tset.disturb[i-1] # change in time
    
    # use variable assignment to create objects to plug into calculations 
  	C_L <- C_L.bio[i-1]*(1-disturbance_severity)
  	C_D <- C_D.bio[i-1] + (C_L.bio[i-1]-C_L)
    M_S <- M_S.bio[i-1]
  	M_C <- M_C.bio[i-1]
  	
  	# calculate change in proportional cover at each timestep
    dC_L <- (dt/5)*(g_C*C_L*(1-C_L-C_D-M_S-M_C) + gamma_C*(1-C_L-C_D-M_S-M_C) - g_M_C_L*C_L*(M_S + M_C) - d_C*C_L) 
    dC_D <- (dt/5)*(d_C*C_L + M_C*(h_M_C + d_M) - g_M*C_D*(M_S + M_C) - e*C_D)
    dM_S <- (dt/5)*(g_M*M_S*(1-C_L-C_D-M_S-M_C) + g_M*M_C*(1-C_L-C_D-M_S-M_C) + gamma_M*(1-C_L-C_D-M_S-M_C) - M_S*(h_M_S + d_M))
    dM_C <- (dt/5)*(g_M*M_C*C_D + g_M*M_S*C_D + g_M_C_L*C_L*(M_S + M_C) + gamma_M*C_D - M_C*(h_M_C + d_M + e))
  	
  	
  	# calculate the proportional cover at each timestep
  	C_L.bio[i] <- C_L + dC_L
  	C_D.bio[i] <- C_D + dC_D
  	M_S.bio[i] <- M_S + dM_S
  	M_C.bio[i] <- M_C + dM_C
    }
  }
# final output is maximum cover for a given severity of disturbance
C_L.bio.eq_cover[j] <- C_L.bio[max(tset.disturb)]
C_D.bio.eq_cover[j] <- C_D.bio[max(tset.disturb)]
M_S.bio.eq_cover[j] <- M_S.bio[max(tset.disturb)]
M_C.bio.eq_cover[j] <- M_C.bio[max(tset.disturb)]
}

```



#### Equilibrium cover of live coral and total macroalgae after 100: high (5 year) frequency

Biological disturbances
```{r Equilibrium cover structure-retaining five years, echo=FALSE}

disturb_set <- seq(from = 0, to = 1, by = .01) # Time vector for outer loop
tset.disturb <- seq(from = 0, to = 1200) # Time vector for inner loop

# create holding vectors for equilibration time following disturbance for all space holders
C_L.bio.eq_cover <- NaN*disturb_set
C_D.bio.eq_cover <- NaN*disturb_set
M_S.bio.eq_cover <- NaN*disturb_set
M_C.bio.eq_cover <- NaN*disturb_set

# create outer for() loop
# for each value j from 1 to the end of disturb_set (so for each value of disturb_set)
for(j in 1:length(disturb_set)){
	# assign value for 'disturb_set' for this iteration Y to that simulation
  disturbance_severity <- disturb_set[j] 
	
  # create holding vectors for the inner loop
	C_L.bio <- NaN*tset.disturb 
	C_L.bio[1] <- 0.6
	C_D.bio <- NaN*tset.disturb 
	C_D.bio[1] <- 0
	M_S.bio <- NaN*tset.disturb 
	M_S.bio[1] <- 0.05
	M_C.bio <- NaN*tset.disturb 
	M_C.bio[1] <- 0.05

# Inner loop	
# for each value of 'disturb_set'
  if(tset_phys_dist_sev[i] !=t_disturb_10 & tset_phys_dist_sev[i] !=t_disturb_11 & tset_phys_dist_sev[i] !=t_disturb_12 & tset_phys_dist_sev[i] !=t_disturb_13 & tset_phys_dist_sev[i] !=t_disturb_14 & tset_phys_dist_sev[i] !=t_disturb_15 & tset_phys_dist_sev[i] !=t_disturb_16 & tset_phys_dist_sev[i] !=t_disturb_17 & tset_phys_dist_sev[i] !=t_disturb_18 & tset_phys_dist_sev[i] !=t_disturb_19 & tset_phys_dist_sev[i] !=t_disturb_20 & tset_phys_dist_sev[i] !=t_disturb_21 & tset_phys_dist_sev[i] !=t_disturb_22 & tset_phys_dist_sev[i] !=t_disturb_23 & tset_phys_dist_sev[i] !=t_disturb_24 & tset_phys_dist_sev[i] !=t_disturb_25 & tset_phys_dist_sev[i] !=t_disturb_26){
    dt <- tset.disturb[i] - tset.disturb[i-1]{ # change in time
    
    # use variable assignment to create objects to plug into calculations 
  	C_L <- C_L.bio[i-1]
  	C_D <- C_D.bio[i-1]
    M_S <- M_S.bio[i-1]
  	M_C <- M_C.bio[i-1]
  	
  	# calculate change in proportional cover at each timestep
    dC_L <- (dt/5)*(g_C*C_L*(1-C_L-C_D-M_S-M_C) + gamma_C*(1-C_L-C_D-M_S-M_C) - g_M_C_L*C_L*(M_S + M_C) - d_C*C_L) 
    dC_D <- (dt/5)*(d_C*C_L + M_C*(h_M_C + d_M) - g_M*C_D*(M_S + M_C) - e*C_D)
    dM_S <- (dt/5)*(g_M*M_S*(1-C_L-C_D-M_S-M_C) + g_M*M_C*(1-C_L-C_D-M_S-M_C) + gamma_M*(1-C_L-C_D-M_S-M_C) - M_S*(h_M_S + d_M))
    dM_C <- (dt/5)*(g_M*M_C*C_D + g_M*M_S*C_D + g_M_C_L*C_L*(M_S + M_C) + gamma_M*C_D - M_C*(h_M_C + d_M + e))
  	
  	
  	# calculate the population density at each timestep
  	C_L.bio[i] <- C_L + dC_L
  	C_D.bio[i] <- C_D + dC_D
  	M_S.bio[i] <- M_S + dM_S
  	M_C.bio[i] <- M_C + dM_C
    }
    
     else {
        dt <- tset.disturb[i] - tset.disturb[i-1] # change in time
    
    # use variable assignment to create objects to plug into calculations 
  	C_L <- C_L.bio[i-1]*(1-disturbance_severity)
  	C_D <- C_D.bio[i-1] + (C_L.bio[i-1]-C_L)
    M_S <- M_S.bio[i-1]
  	M_C <- M_C.bio[i-1]
  	
  	# calculate change in proportional cover at each timestep
    dC_L <- (dt/5)*(g_C*C_L*(1-C_L-C_D-M_S-M_C) + gamma_C*(1-C_L-C_D-M_S-M_C) - g_M_C_L*C_L*(M_S + M_C) - d_C*C_L) 
    dC_D <- (dt/5)*(d_C*C_L + M_C*(h_M_C + d_M) - g_M*C_D*(M_S + M_C) - e*C_D)
    dM_S <- (dt/5)*(g_M*M_S*(1-C_L-C_D-M_S-M_C) + g_M*M_C*(1-C_L-C_D-M_S-M_C) + gamma_M*(1-C_L-C_D-M_S-M_C) - M_S*(h_M_S + d_M))
    dM_C <- (dt/5)*(g_M*M_C*C_D + g_M*M_S*C_D + g_M_C_L*C_L*(M_S + M_C) + gamma_M*C_D - M_C*(h_M_C + d_M + e))
  	
  	
  	# calculate the proportional cover at each timestep
  	C_L.bio[i] <- C_L + dC_L
  	C_D.bio[i] <- C_D + dC_D
  	M_S.bio[i] <- M_S + dM_S
  	M_C.bio[i] <- M_C + dM_C
    }
  }
# final output is maximum cover for a given severity of disturbance
C_L.bio.eq_cover[j] <- C_L.bio[max(tset.disturb)]
C_D.bio.eq_cover[j] <- C_D.bio[max(tset.disturb)]
M_S.bio.eq_cover[j] <- M_S.bio[max(tset.disturb)]
M_C.bio.eq_cover[j] <- M_C.bio[max(tset.disturb)]
}

```

Equilibrium cover after structure-removing disturbances every 5 years
```{r Equilibrium cover structure-removing five years, echo=FALSE}

disturb_set <- seq(from = 0, to = 1, by = .01) # Time vector for outer loop
tset.disturb <- seq(from = 0, to = 1200) # Time vector for inner loop

# create holding vectors for equilibration time following disturbance for all space holders
C_L.phys.eq_cover <- NaN*disturb_set
C_D.phys.eq_cover <- NaN*disturb_set
M_S.phys.eq_cover <- NaN*disturb_set
M_C.phys.eq_cover <- NaN*disturb_set

# create outer for() loop
# for each value j from 1 to the end of disturb_set (so for each value of disturb_set)
for(j in 1:length(disturb_set)){
	# assign value for 'disturb_set' for this iteration Y to that simulation
  disturbance_severity <- disturb_set[j] 
	
  # create holding vectors for the inner loop
	C_L.phys <- NaN*tset.disturb 
	C_L.phys[1] <- 0.6
	C_D.phys <- NaN*tset.disturb 
	C_D.phys[1] <- 0
	M_S.phys <- NaN*tset.disturb 
	M_S.phys[1] <- 0.05
	M_C.phys <- NaN*tset.disturb 
	M_C.phys[1] <- 0.05

# Inner loop	
# for each value of 'disturb_set'
  for(i in 2:length(tset.disturb)){
  	
    if(tset.disturb[i] !=t_disturb_1 & tset.disturb[i] !=t_disturb_2 & tset.disturb[i] !=t_disturb_3){
    dt <- tset.disturb[i] - tset.disturb[i-1] # change in time
    
    # use variable assignment to create objects to plug into calculations 
  	C_L <- C_L.phys[i-1]
  	C_D <- C_D.phys[i-1]
    M_S <- M_S.phys[i-1]
  	M_C <- M_C.phys[i-1]
  	
  	# calculate change in proportional cover at each timestep
    dC_L <- (dt/5)*(g_C*C_L*(1-C_L-C_D-M_S-M_C) + gamma_C*(1-C_L-C_D-M_S-M_C) - g_M_C_L*C_L*(M_S + M_C) - d_C*C_L) 
    dC_D <- (dt/5)*(d_C*C_L + M_C*(h_M_C + d_M) - g_M*C_D*(M_S + M_C) - e*C_D)
    dM_S <- (dt/5)*(g_M*M_S*(1-C_L-C_D-M_S-M_C) + g_M*M_C*(1-C_L-C_D-M_S-M_C) + gamma_M*(1-C_L-C_D-M_S-M_C) - M_S*(h_M_S + d_M))
    dM_C <- (dt/5)*(g_M*M_C*C_D + g_M*M_S*C_D + g_M_C_L*C_L*(M_S + M_C) + gamma_M*C_D - M_C*(h_M_C + d_M + e))
  	
  	
  	# calculate the population density at each timestep
  	C_L.phys[i] <- C_L + dC_L
  	C_D.phys[i] <- C_D + dC_D
  	M_S.phys[i] <- M_S + dM_S
  	M_C.phys[i] <- M_C + dM_C
    }
    
     else {
        dt <- tset.disturb[i] - tset.disturb[i-1] # change in time
    
    # use variable assignment to create objects to plug into calculations 
  	C_L <- C_L.phys[i-1]*(1-disturbance_severity)
  	C_D <- C_D.phys[i-1]*(1-disturbance_severity)
    M_S <- M_S.phys[i-1]*(1-disturbance_severity)
  	M_C <- M_C.phys[i-1]*(1-disturbance_severity)
  	
  	# calculate change in proportional cover at each timestep
    dC_L <- (dt/5)*(g_C*C_L*(1-C_L-C_D-M_S-M_C) + gamma_C*(1-C_L-C_D-M_S-M_C) - g_M_C_L*C_L*(M_S + M_C) - d_C*C_L) 
    dC_D <- (dt/5)*(d_C*C_L + M_C*(h_M_C + d_M) - g_M*C_D*(M_S + M_C) - e*C_D)
    dM_S <- (dt/5)*(g_M*M_S*(1-C_L-C_D-M_S-M_C) + g_M*M_C*(1-C_L-C_D-M_S-M_C) + gamma_M*(1-C_L-C_D-M_S-M_C) - M_S*(h_M_S + d_M))
    dM_C <- (dt/5)*(g_M*M_C*C_D + g_M*M_S*C_D + g_M_C_L*C_L*(M_S + M_C) + gamma_M*C_D - M_C*(h_M_C + d_M + e))
  	
  	
  	# calculate the proportional cover at each timestep
  	C_L.phys[i] <- C_L + dC_L
  	C_D.phys[i] <- C_D + dC_D
  	M_S.phys[i] <- M_S + dM_S
  	M_C.phys[i] <- M_C + dM_C
    }
  }
# final output is equilibrium cover for a given severity of disturbance
C_L.phys.eq_cover[j] <- C_L.phys[max(tset.disturb)]
C_D.phys.eq_cover[j] <- C_D.phys[max(tset.disturb)]
M_S.phys.eq_cover[j] <- M_S.phys[max(tset.disturb)]
M_C.phys.eq_cover[j] <- M_C.phys[max(tset.disturb)]
}

```

Equilibrium cover, 5 years
```{r Equilibrium cover plots five years, echo=FALSE}

# Live coral
plot(disturb_set, C_L.phys.eq_cover.half, col = "blue", lwd = 2, lty = 'solid', type = 'l', xlab = "Severity of disturbance", ylab = "Equilibrium cover", xlim = c(0, 1), ylim = c(0, 1.1), main = "Live coral")
lines(disturb_set, C_L.bio.eq_cover.half, col = "red", lwd = 2, lty = 'solid', type = 'l')
#lines(disturb_set, C_L.phys.eq_cover.two_thirds, col = "blue", lwd = 1, lty = 'dashed', type = 'l')
#lines(disturb_set, C_L.bio.eq_cover.two_thirds, col = "red", lwd = 1, lty = 'dashed', type = 'l')
#lines(disturb_set, C_L.phys.eq_cover.third, col = "red", lwd = 1, lty = 'dashed', type = 'l')
#lines(disturb_set, C_L.bio.eq_cover.third, col = "red", lwd = 1, lty = 'dashed', type = 'l')
legend(x=0.6,y=1.1,legend=c('Physical disturbance','Biological disturbance'),lwd=1,lty='solid',col=c('blue','red'))
#legend(x=.6, y = 0.8, legend=c('hmc = 1/2(hms)', 'hmc = 1/3(hms)'),lwd=1,lty=c('dotted', 'solid', 'dashed'),col='black')

# Total macroalgae
plot(disturb_set, c(M_C.phys.eq_cover.half + M_S.phys.eq_cover.half), col = "blue", lwd = 2, lty = 'solid', type = 'l', xlab = "Severity of disturbance", ylab = "Equilibrium cover", xlim = c(0, 1), ylim = c(0, 1.1), main = "Total macroalgae")
lines(disturb_set, c(M_C.bio.eq_cover.half + M_S.bio.eq_cover.half), col = "red", lwd = 2, lty = 'solid', type = 'l')
#lines(disturb_set, c(M_C.phys.eq_cover.two_thirds + M_S.phys.eq_cover.two_thirds), col = "blue", lwd = 1, lty = 'dashed', type = 'l')
#lines(disturb_set, c(M_C.bio.eq_cover.two_thirds + M_S.bio.eq_cover.two_thirds), col = "red", lwd = 1, lty = 'dashed', type = 'l')
#lines(disturb_set, c(M_C.phys.eq_cover.third + M_S.phys.eq_cover.third), col = "blue", lwd = 1, lty = 'dashed', type = 'l')
#lines(disturb_set, c(M_C.bio.eq_cover.third + M_S.bio.eq_cover.third), col = "red", lwd = 1, lty = 'dashed', type = 'l')
legend(x=0.6,y=1.1,legend=c('Structure-removing','Structure-retaining'), lwd=1, lty='solid', col=c('blue','red'))
#legend(x=.6, y = 0.8, legend=c('Grazing pressure = 1/2', 'Grazing pressure = 1/3'),lwd=1,lty=c('solid', 'dashed'),col='black')

```

### Disturbance sequences: 

#### Time series: 50% disturbances in different sequences

Structure removing first, five years apart
```{r 50% structure-removing first five years apart, echo = FALSE}

# Create time vector, holding vectors for each state variable, and set initial conditions
tset_phys_dist_mult <- seq(from = 0, to = 259)
C_L.simu_phys_mult <- NaN*tset_phys_dist_mult; C_L.simu_phys_mult[1] <- 0.6
C_D.simu_phys_mult <- NaN*tset_phys_dist_mult; C_D.simu_phys_mult[1] <- 0
M_S.simu_phys_mult <- NaN*tset_phys_dist_mult; M_S.simu_phys_mult[1] <- 0.05
M_C.simu_phys_mult <- NaN*tset_phys_dist_mult; M_C.simu_phys_mult[1] <- 0.05

# For() loop with if/else statement to simulate trajectories of spatial occupancy by live coral and macroalgae before and after disturbance

for(i in 2:length(tset_phys_dist_mult)){ # For every element after the first in the time vector...
  
  if(tset_phys_dist_mult[i] !=t_disturb){ # ...if the time point does not equal the time of disturbance, run the simulation as normal...
  
  dt <- tset_phys_dist_mult[i] - tset_phys_dist_mult[i-1] # calculate change in time  
 
  # Proportional cover at previous time step  
  C_L <- C_L.simu_phys_mult[i-1]
  C_D <- C_D.simu_phys_mult[i-1]
  M_S <- M_S.simu_phys_mult[i-1]
  M_C <- M_C.simu_phys_mult[i-1]
  
  # Change in proportional cover, converted to units of months (/5)
  dC_L <- (dt/5)*(g_C*C_L*(1-C_L-C_D-M_S-M_C) + gamma_C*(1-C_L-C_D-M_S-M_C) - g_M_C_L*C_L*(M_S + M_C) - d_C*C_L) 
  dC_D <- (dt/5)*(d_C*C_L + M_C*(h_M_C + d_M) - g_M*C_D*(M_S + M_C) - e*C_D)
  dM_S <- (dt/5)*(g_M*M_S*(1-C_L-C_D-M_S-M_C) + g_M*M_C*(1-C_L-C_D-M_S-M_C) + gamma_M*(1-C_L-C_D-M_S-M_C) - M_S*(h_M_S + d_M))
  dM_C <- (dt/5)*(g_M*M_C*C_D + g_M*M_S*C_D + g_M_C_L*C_L*(M_S + M_C) + gamma_M*C_D - M_C*(h_M_C + d_M + e))
 
  # Proportional cover at new time step   
  C_L.simu_phys_mult[i] <- C_L.simu_phys_mult[i-1] + dC_L
  C_D.simu_phys_mult[i] <- C_D.simu_phys_mult[i-1] + dC_D
  M_S.simu_phys_mult[i] <- M_S.simu_phys_mult[i-1] + dM_S
  M_C.simu_phys_mult[i] <- M_C.simu_phys_mult[i-1] + dM_C
  }
  
    else { # ... but if the time step does equal the time of disturbance...
  
      dt <- tset_phys_dist_mult[i] - tset_phys_dist_mult[i-1]

  # reduce all space holders to 50% of their pre-disturbance levels
  C_L <- C_L.simu_phys_mult[i-1]*0.5
  C_D <- C_D.simu_phys_mult[i-1]*0.5
  M_S <- M_S.simu_phys_mult[i-1]*0.5
  M_C <- M_C.simu_phys_mult[i-1]*0.5
  
  dC_L <- (dt/5)*(g_C*C_L*(1-C_L-C_D-M_S-M_C) + gamma_C*(1-C_L-C_D-M_S-M_C) - g_M_C_L*C_L*(M_S + M_C) - d_C*C_L) 
  dC_D <- (dt/5)*(d_C*C_L + M_C*(h_M_C + d_M) - g_M*C_D*(M_S + M_C) - e*C_D)
  dM_S <- (dt/5)*(g_M*M_S*(1-C_L-C_D-M_S-M_C) + g_M*M_C*(1-C_L-C_D-M_S-M_C) + gamma_M*(1-C_L-C_D-M_S-M_C) - M_S*(h_M_S + d_M))
  dM_C <- (dt/5)*(g_M*M_C*C_D + g_M*M_S*C_D + g_M_C_L*C_L*(M_S + M_C) + gamma_M*C_D - M_C*(h_M_C + d_M + e))
  
  C_L.simu_phys_mult[i] <- C_L+ dC_L
  C_D.simu_phys_mult[i] <- C_D+ dC_D
  M_S.simu_phys_mult[i] <- M_S+ dM_S
  M_C.simu_phys_mult[i] <- M_C+ dM_C
  }
  
}

```

 
```{r 50% structure-retaining second five years apart, echo = FALSE}

# Create time vector, holding vectors for each state variable, and set initial conditions
tset_bio_dist_mult <- seq(from = 259, to = 1200)
C_L.bio <- NaN*tset_bio_dist_mult; C_L.bio[1] <- C_L.simu_phys_mult[260]
C_D.bio <- NaN*tset_bio_dist_mult; C_D.bio[1] <- C_D.simu_phys_mult[260]
M_S.bio <- NaN*tset_bio_dist_mult; M_S.bio[1] <- M_S.simu_phys_mult[260]
M_C.bio <- NaN*tset_bio_dist_mult; M_C.bio[1] <- M_C.simu_phys_mult[260]

# For() loop with if/else statement to simulate trajectories of spatial occupancy by live coral and macroalgae before and after disturbance
for(i in 2:length(tset_bio_dist_mult)){ # For every element after the first in the time vector...
  
  if(tset_bio_dist_mult[i] !=260){ # ...if the time point does not equal the time of disturbance, run the simulation as normal...

# calculate change in time  
  dt <- tset_bio_dist_mult[i] - tset_bio_dist_mult[i-1] 

  # Proportional cover at previous time step  
  C_L <- C_L.bio[i-1] 
  C_D <- C_D.bio[i-1]
  M_S <- M_S.bio[i-1]
  M_C <- M_C.bio[i-1]
 
# Change in proportional cover, converted to units of months (/5)
  dC_L <- (dt/5)*(g_C*C_L*(1-C_L-C_D-M_S-M_C) + gamma_C*(1-C_L-C_D-M_S-M_C) - g_M_C_L*C_L*(M_S + M_C) - d_C*C_L) 
  dC_D <- (dt/5)*(d_C*C_L + M_C*(h_M_C + d_M) - g_M*C_D*(M_S + M_C) - e*C_D)
  dM_S <- (dt/5)*(g_M*M_S*(1-C_L-C_D-M_S-M_C) + g_M*M_C*(1-C_L-C_D-M_S-M_C) + gamma_M*(1-C_L-C_D-M_S-M_C) - M_S*(h_M_S + d_M))
  dM_C <- (dt/5)*(g_M*M_C*C_D + g_M*M_S*C_D + g_M_C_L*C_L*(M_S + M_C) + gamma_M*C_D - M_C*(h_M_C + d_M + e))

# Proportional cover at new time step  
  C_L.bio[i] <- C_L.bio[i-1] + dC_L
  C_D.bio[i] <- C_D.bio[i-1] + dC_D
  M_S.bio[i] <- M_S.bio[i-1] + dM_S
  M_C.bio[i] <- M_C.bio[i-1] + dM_C
  
  }

  # ... but if the time step does equal the time of disturbance...
    else {
  
  dt <- tset_bio_dist_mult[i] - tset_bio_dist_mult[i-1]

  C_L <- C_L.bio[i-1]*0.5 # reduce live coral to 50% of its pre-disturbance level, and transfer this directly to the dead coral class
  C_D <- C_D.bio[i-1] + (C_L.bio[i-1]-C_L)
  M_S <- M_S.bio[i-1]
  M_C <- M_C.bio[i-1]
  
  dC_L <- (dt/5)*(g_C*C_L*(1-C_L-C_D-M_S-M_C) + gamma_C*(1-C_L-C_D-M_S-M_C) - g_M_C_L*C_L*(M_S + M_C) - d_C*C_L) 
  dC_D <- (dt/5)*(d_C*C_L + M_C*(h_M_C + d_M) - g_M*C_D*(M_S + M_C) - e*C_D)
  dM_S <- (dt/5)*(g_M*M_S*(1-C_L-C_D-M_S-M_C) + g_M*M_C*(1-C_L-C_D-M_S-M_C) + gamma_M*(1-C_L-C_D-M_S-M_C) - M_S*(h_M_S + d_M))
  dM_C <- (dt/5)*(g_M*M_C*C_D + g_M*M_S*C_D + g_M_C_L*C_L*(M_S + M_C) + gamma_M*C_D - M_C*(h_M_C + d_M + e))
  
  C_L.bio[i] <- C_L + dC_L
  C_D.bio[i] <- C_D + dC_D
  M_S.bio[i] <- M_S + dM_S
  M_C.bio[i] <- M_C + dM_C
  
  }
  
}

```

Plot of time series for structure removing followed by structure retaining disturbance, 5 years apart
```{r plot of structure-removing first five years apart}
plot(tset_phys_dist_mult, C_L.simu_phys_mult, col = "blue", lwd = 1, lty = 'solid', type = 'l', xlab = "Time (months)", ylab = "Proportional cover", xlim = c(0, max(tset_bio_dist_mult)), ylim = c(0, 1), main = "Structure-removing")
lines(tset_phys_dist_mult, C_D.simu_phys_mult, col = "black", lwd = 1, lty = 'solid')
lines(tset_phys_dist_mult, M_S.simu_phys_mult, col = "green", lwd = 1, lty = 'solid')
lines(tset_phys_dist_mult, M_C.simu_phys_mult, col = "grey", lwd = 1, lty = 'solid')
lines(tset_bio_dist_mult, C_L.bio, col = "blue", lwd = 1, lty = 'solid')
lines(tset_bio_dist_mult, C_D.bio, col = "black", lwd = 1, lty = 'solid')
lines(tset_bio_dist_mult, M_S.bio, col = "green", lwd = 1, lty = 'solid')
lines(tset_bio_dist_mult, M_C.bio, col = "grey", lwd = 1, lty = 'solid')
abline(v=200, lty = 'dashed', col = "blue")
abline(v=260, lty = 'dashed', col = "red")
```

Structure retaining first, 5 years
```{r 50% structure-retaining first 5 years apart, echo = FALSE}

# Create time vector, holding vectors for each state variable, and set initial conditions
tset_bio_dist_mult <- seq(from = 0, to = 259)
C_L.bio <- NaN*tset_bio_dist_mult; C_L.bio[1] <- 0.6
C_D.bio <- NaN*tset_bio_dist_mult; C_D.bio[1] <- 0
M_S.bio <- NaN*tset_bio_dist_mult; M_S.bio[1] <- 0.05
M_C.bio <- NaN*tset_bio_dist_mult; M_C.bio[1] <- 0.05

# For() loop with if/else statement to simulate trajectories of spatial occupancy by live coral and macroalgae before and after disturbance
for(i in 2:length(tset_bio_dist_mult)){ # For every element after the first in the time vector...
  
  if(tset_bio_dist_mult[i] !=200){ # ...if the time point does not equal the time of disturbance, run the simulation as normal...

# calculate change in time  
  dt <- tset_bio_dist_mult[i] - tset_bio_dist_mult[i-1] 

  # Proportional cover at previous time step  
  C_L <- C_L.bio[i-1] 
  C_D <- C_D.bio[i-1]
  M_S <- M_S.bio[i-1]
  M_C <- M_C.bio[i-1]
 
# Change in proportional cover, converted to units of months (/5)
  dC_L <- (dt/5)*(g_C*C_L*(1-C_L-C_D-M_S-M_C) + gamma_C*(1-C_L-C_D-M_S-M_C) - g_M_C_L*C_L*(M_S + M_C) - d_C*C_L) 
  dC_D <- (dt/5)*(d_C*C_L + M_C*(h_M_C + d_M) - g_M*C_D*(M_S + M_C) - e*C_D)
  dM_S <- (dt/5)*(g_M*M_S*(1-C_L-C_D-M_S-M_C) + g_M*M_C*(1-C_L-C_D-M_S-M_C) + gamma_M*(1-C_L-C_D-M_S-M_C) - M_S*(h_M_S + d_M))
  dM_C <- (dt/5)*(g_M*M_C*C_D + g_M*M_S*C_D + g_M_C_L*C_L*(M_S + M_C) + gamma_M*C_D - M_C*(h_M_C + d_M + e))

# Proportional cover at new time step  
  C_L.bio[i] <- C_L.bio[i-1] + dC_L
  C_D.bio[i] <- C_D.bio[i-1] + dC_D
  M_S.bio[i] <- M_S.bio[i-1] + dM_S
  M_C.bio[i] <- M_C.bio[i-1] + dM_C
  
  }

  # ... but if the time step does equal the time of disturbance...
    else {
  
  dt <- tset_bio_dist_mult[i] - tset_bio_dist_mult[i-1]

  C_L <- C_L.bio[i-1]*0.5 # reduce live coral to 50% of its pre-disturbance level, and transfer this directly to the dead coral class
  C_D <- C_D.bio[i-1] + (C_L.bio[i-1]-C_L)
  M_S <- M_S.bio[i-1]
  M_C <- M_C.bio[i-1]
  
  dC_L <- (dt/5)*(g_C*C_L*(1-C_L-C_D-M_S-M_C) + gamma_C*(1-C_L-C_D-M_S-M_C) - g_M_C_L*C_L*(M_S + M_C) - d_C*C_L) 
  dC_D <- (dt/5)*(d_C*C_L + M_C*(h_M_C + d_M) - g_M*C_D*(M_S + M_C) - e*C_D)
  dM_S <- (dt/5)*(g_M*M_S*(1-C_L-C_D-M_S-M_C) + g_M*M_C*(1-C_L-C_D-M_S-M_C) + gamma_M*(1-C_L-C_D-M_S-M_C) - M_S*(h_M_S + d_M))
  dM_C <- (dt/5)*(g_M*M_C*C_D + g_M*M_S*C_D + g_M_C_L*C_L*(M_S + M_C) + gamma_M*C_D - M_C*(h_M_C + d_M + e))
  
  C_L.bio[i] <- C_L + dC_L
  C_D.bio[i] <- C_D + dC_D
  M_S.bio[i] <- M_S + dM_S
  M_C.bio[i] <- M_C + dM_C
  
  }
  
}
```


```{r 50% structure-removing second 5 years apart, echo = FALSE}

# Create time vector, holding vectors for each state variable, and set initial conditions
tset_phys_dist_mult <- seq(from = 259, to = 1200)
C_L.simu_phys_mult <- NaN*tset_phys_dist_mult; C_L.simu_phys_mult[1] <- C_L.bio[260]
C_D.simu_phys_mult <- NaN*tset_phys_dist_mult; C_D.simu_phys_mult[1] <- C_D.bio[260]
M_S.simu_phys_mult <- NaN*tset_phys_dist_mult; M_S.simu_phys_mult[1] <- M_S.bio[260]
M_C.simu_phys_mult <- NaN*tset_phys_dist_mult; M_C.simu_phys_mult[1] <- M_C.bio[260]


# For() loop with if/else statement to simulate trajectories of spatial occupancy by live coral and macroalgae before and after disturbance

for(i in 2:length(tset_phys_dist_mult)){ # For every element after the first in the time vector...
  
  if(tset_phys_dist_mult[i] !=260){ # ...if the time point does not equal the time of disturbance, run the simulation as normal...
  
  dt <- tset_phys_dist_mult[i] - tset_phys_dist_mult[i-1] # calculate change in time  
 
  # Proportional cover at previous time step  
  C_L <- C_L.simu_phys_mult[i-1]
  C_D <- C_D.simu_phys_mult[i-1]
  M_S <- M_S.simu_phys_mult[i-1]
  M_C <- M_C.simu_phys_mult[i-1]
  
  # Change in proportional cover, converted to units of months (/5)
  dC_L <- (dt/5)*(g_C*C_L*(1-C_L-C_D-M_S-M_C) + gamma_C*(1-C_L-C_D-M_S-M_C) - g_M_C_L*C_L*(M_S + M_C) - d_C*C_L) 
  dC_D <- (dt/5)*(d_C*C_L + M_C*(h_M_C + d_M) - g_M*C_D*(M_S + M_C) - e*C_D)
  dM_S <- (dt/5)*(g_M*M_S*(1-C_L-C_D-M_S-M_C) + g_M*M_C*(1-C_L-C_D-M_S-M_C) + gamma_M*(1-C_L-C_D-M_S-M_C) - M_S*(h_M_S + d_M))
  dM_C <- (dt/5)*(g_M*M_C*C_D + g_M*M_S*C_D + g_M_C_L*C_L*(M_S + M_C) + gamma_M*C_D - M_C*(h_M_C + d_M + e))
 
  # Proportional cover at new time step   
  C_L.simu_phys_mult[i] <- C_L.simu_phys_mult[i-1] + dC_L
  C_D.simu_phys_mult[i] <- C_D.simu_phys_mult[i-1] + dC_D
  M_S.simu_phys_mult[i] <- M_S.simu_phys_mult[i-1] + dM_S
  M_C.simu_phys_mult[i] <- M_C.simu_phys_mult[i-1] + dM_C
  }
  
    else { # ... but if the time step does equal the time of disturbance...
  
      dt <- tset_phys_dist_mult[i] - tset_phys_dist_mult[i-1]

  # reduce all space holders to 50% of their pre-disturbance levels
  C_L <- C_L.simu_phys_mult[i-1]*0.5
  C_D <- C_D.simu_phys_mult[i-1]*0.5
  M_S <- M_S.simu_phys_mult[i-1]*0.5
  M_C <- M_C.simu_phys_mult[i-1]*0.5
  
  dC_L <- (dt/5)*(g_C*C_L*(1-C_L-C_D-M_S-M_C) + gamma_C*(1-C_L-C_D-M_S-M_C) - g_M_C_L*C_L*(M_S + M_C) - d_C*C_L) 
  dC_D <- (dt/5)*(d_C*C_L + M_C*(h_M_C + d_M) - g_M*C_D*(M_S + M_C) - e*C_D)
  dM_S <- (dt/5)*(g_M*M_S*(1-C_L-C_D-M_S-M_C) + g_M*M_C*(1-C_L-C_D-M_S-M_C) + gamma_M*(1-C_L-C_D-M_S-M_C) - M_S*(h_M_S + d_M))
  dM_C <- (dt/5)*(g_M*M_C*C_D + g_M*M_S*C_D + g_M_C_L*C_L*(M_S + M_C) + gamma_M*C_D - M_C*(h_M_C + d_M + e))
  
  C_L.simu_phys_mult[i] <- C_L+ dC_L
  C_D.simu_phys_mult[i] <- C_D+ dC_D
  M_S.simu_phys_mult[i] <- M_S+ dM_S
  M_C.simu_phys_mult[i] <- M_C+ dM_C
  }
  
}
```

Plot of time series for structure removing followed by structure retaining disturbance
```{r plot of 50% structure retaining first five years apart}

plot(tset_phys_dist_mult, C_L.simu_phys_mult, col = "blue", lwd = 1, lty = 'solid', type = 'l', xlab = "Time (months)", ylab = "Proportional cover", xlim = c(0, max(tset_phys_dist_mult)), ylim = c(0, 1), main = "Structure-removing")
lines(tset_phys_dist_mult, C_D.simu_phys_mult, col = "black", lwd = 1, lty = 'solid')
lines(tset_phys_dist_mult, M_S.simu_phys_mult, col = "green", lwd = 1, lty = 'solid')
lines(tset_phys_dist_mult, M_C.simu_phys_mult, col = "grey", lwd = 1, lty = 'solid')
lines(tset_bio_dist_mult, C_L.bio, col = "blue", lwd = 1, lty = 'solid')
lines(tset_bio_dist_mult, C_D.bio, col = "black", lwd = 1, lty = 'solid')
lines(tset_bio_dist_mult, M_S.bio, col = "green", lwd = 1, lty = 'solid')
lines(tset_bio_dist_mult, M_C.bio, col = "grey", lwd = 1, lty = 'solid')
abline(v=200, lty = 'dashed', col = "blue")
abline(v=260, lty = 'dashed', col = "red")
```


Structure removing first, ten years apart (The Oli-bleaching scenario)
```{r 50% structure-removing first ten years apart, echo = FALSE}

# Create time vector, holding vectors for each state variable, and set initial conditions
tset_phys_dist_mult <- seq(from = 0, to = 319)
C_L.simu_phys_mult <- NaN*tset_phys_dist_mult; C_L.simu_phys_mult[1] <- 0.6
C_D.simu_phys_mult <- NaN*tset_phys_dist_mult; C_D.simu_phys_mult[1] <- 0
M_S.simu_phys_mult <- NaN*tset_phys_dist_mult; M_S.simu_phys_mult[1] <- 0.05
M_C.simu_phys_mult <- NaN*tset_phys_dist_mult; M_C.simu_phys_mult[1] <- 0.05

# For() loop with if/else statement to simulate trajectories of spatial occupancy by live coral and macroalgae before and after disturbance

for(i in 2:length(tset_phys_dist_mult)){ # For every element after the first in the time vector...
  
  if(tset_phys_dist_mult[i] !=200){ # ...if the time point does not equal the time of disturbance, run the simulation as normal...
  
  dt <- tset_phys_dist_mult[i] - tset_phys_dist_mult[i-1] # calculate change in time  
 
  # Proportional cover at previous time step  
  C_L <- C_L.simu_phys_mult[i-1]
  C_D <- C_D.simu_phys_mult[i-1]
  M_S <- M_S.simu_phys_mult[i-1]
  M_C <- M_C.simu_phys_mult[i-1]
  
  # Change in proportional cover, converted to units of months (/5)
  dC_L <- (dt/5)*(g_C*C_L*(1-C_L-C_D-M_S-M_C) + gamma_C*(1-C_L-C_D-M_S-M_C) - g_M_C_L*C_L*(M_S + M_C) - d_C*C_L) 
  dC_D <- (dt/5)*(d_C*C_L + M_C*(h_M_C + d_M) - g_M*C_D*(M_S + M_C) - e*C_D)
  dM_S <- (dt/5)*(g_M*M_S*(1-C_L-C_D-M_S-M_C) + g_M*M_C*(1-C_L-C_D-M_S-M_C) + gamma_M*(1-C_L-C_D-M_S-M_C) - M_S*(h_M_S + d_M))
  dM_C <- (dt/5)*(g_M*M_C*C_D + g_M*M_S*C_D + g_M_C_L*C_L*(M_S + M_C) + gamma_M*C_D - M_C*(h_M_C + d_M + e))
 
  # Proportional cover at new time step   
  C_L.simu_phys_mult[i] <- C_L.simu_phys_mult[i-1] + dC_L
  C_D.simu_phys_mult[i] <- C_D.simu_phys_mult[i-1] + dC_D
  M_S.simu_phys_mult[i] <- M_S.simu_phys_mult[i-1] + dM_S
  M_C.simu_phys_mult[i] <- M_C.simu_phys_mult[i-1] + dM_C
  }
  
    else { # ... but if the time step does equal the time of disturbance...
  
      dt <- tset_phys_dist_mult[i] - tset_phys_dist_mult[i-1]

  # reduce all space holders to 50% of their pre-disturbance levels
  C_L <- C_L.simu_phys_mult[i-1]*0.5
  C_D <- C_D.simu_phys_mult[i-1]*0.5
  M_S <- M_S.simu_phys_mult[i-1]*0.5
  M_C <- M_C.simu_phys_mult[i-1]*0.5
  
  dC_L <- (dt/5)*(g_C*C_L*(1-C_L-C_D-M_S-M_C) + gamma_C*(1-C_L-C_D-M_S-M_C) - g_M_C_L*C_L*(M_S + M_C) - d_C*C_L) 
  dC_D <- (dt/5)*(d_C*C_L + M_C*(h_M_C + d_M) - g_M*C_D*(M_S + M_C) - e*C_D)
  dM_S <- (dt/5)*(g_M*M_S*(1-C_L-C_D-M_S-M_C) + g_M*M_C*(1-C_L-C_D-M_S-M_C) + gamma_M*(1-C_L-C_D-M_S-M_C) - M_S*(h_M_S + d_M))
  dM_C <- (dt/5)*(g_M*M_C*C_D + g_M*M_S*C_D + g_M_C_L*C_L*(M_S + M_C) + gamma_M*C_D - M_C*(h_M_C + d_M + e))
  
  C_L.simu_phys_mult[i] <- C_L+ dC_L
  C_D.simu_phys_mult[i] <- C_D+ dC_D
  M_S.simu_phys_mult[i] <- M_S+ dM_S
  M_C.simu_phys_mult[i] <- M_C+ dM_C
  }
  
}

```


```{r 50% structure-retaining second ten years apart, echo = FALSE}

# Create time vector, holding vectors for each state variable, and set initial conditions
tset_bio_dist_mult <- seq(from = 319, to = 1200)
C_L.bio <- NaN*tset_bio_dist_mult; C_L.bio[1] <- C_L.simu_phys_mult[320]
C_D.bio <- NaN*tset_bio_dist_mult; C_D.bio[1] <- C_D.simu_phys_mult[320]
M_S.bio <- NaN*tset_bio_dist_mult; M_S.bio[1] <- M_S.simu_phys_mult[320]
M_C.bio <- NaN*tset_bio_dist_mult; M_C.bio[1] <- M_C.simu_phys_mult[320]

# For() loop with if/else statement to simulate trajectories of spatial occupancy by live coral and macroalgae before and after disturbance
for(i in 2:length(tset_bio_dist_mult)){ # For every element after the first in the time vector...
  
  if(tset_bio_dist_mult[i] !=320){ # ...if the time point does not equal the time of disturbance, run the simulation as normal...

# calculate change in time  
  dt <- tset_bio_dist_mult[i] - tset_bio_dist_mult[i-1] 

  # Proportional cover at previous time step  
  C_L <- C_L.bio[i-1] 
  C_D <- C_D.bio[i-1]
  M_S <- M_S.bio[i-1]
  M_C <- M_C.bio[i-1]
 
# Change in proportional cover, converted to units of months (/5)
  dC_L <- (dt/5)*(g_C*C_L*(1-C_L-C_D-M_S-M_C) + gamma_C*(1-C_L-C_D-M_S-M_C) - g_M_C_L*C_L*(M_S + M_C) - d_C*C_L) 
  dC_D <- (dt/5)*(d_C*C_L + M_C*(h_M_C + d_M) - g_M*C_D*(M_S + M_C) - e*C_D)
  dM_S <- (dt/5)*(g_M*M_S*(1-C_L-C_D-M_S-M_C) + g_M*M_C*(1-C_L-C_D-M_S-M_C) + gamma_M*(1-C_L-C_D-M_S-M_C) - M_S*(h_M_S + d_M))
  dM_C <- (dt/5)*(g_M*M_C*C_D + g_M*M_S*C_D + g_M_C_L*C_L*(M_S + M_C) + gamma_M*C_D - M_C*(h_M_C + d_M + e))

# Proportional cover at new time step  
  C_L.bio[i] <- C_L.bio[i-1] + dC_L
  C_D.bio[i] <- C_D.bio[i-1] + dC_D
  M_S.bio[i] <- M_S.bio[i-1] + dM_S
  M_C.bio[i] <- M_C.bio[i-1] + dM_C
  
  }

  # ... but if the time step does equal the time of disturbance...
    else {
  
  dt <- tset_bio_dist_mult[i] - tset_bio_dist_mult[i-1]

  C_L <- C_L.bio[i-1]*0.5 # reduce live coral to 50% of its pre-disturbance level, and transfer this directly to the dead coral class
  C_D <- C_D.bio[i-1] + (C_L.bio[i-1]-C_L)
  M_S <- M_S.bio[i-1]
  M_C <- M_C.bio[i-1]
  
  dC_L <- (dt/5)*(g_C*C_L*(1-C_L-C_D-M_S-M_C) + gamma_C*(1-C_L-C_D-M_S-M_C) - g_M_C_L*C_L*(M_S + M_C) - d_C*C_L) 
  dC_D <- (dt/5)*(d_C*C_L + M_C*(h_M_C + d_M) - g_M*C_D*(M_S + M_C) - e*C_D)
  dM_S <- (dt/5)*(g_M*M_S*(1-C_L-C_D-M_S-M_C) + g_M*M_C*(1-C_L-C_D-M_S-M_C) + gamma_M*(1-C_L-C_D-M_S-M_C) - M_S*(h_M_S + d_M))
  dM_C <- (dt/5)*(g_M*M_C*C_D + g_M*M_S*C_D + g_M_C_L*C_L*(M_S + M_C) + gamma_M*C_D - M_C*(h_M_C + d_M + e))
  
  C_L.bio[i] <- C_L + dC_L
  C_D.bio[i] <- C_D + dC_D
  M_S.bio[i] <- M_S + dM_S
  M_C.bio[i] <- M_C + dM_C
  
  }
  
}

```

Plot of time series for structure removing followed by structure retaining disturbance, ten years apart
```{r structure-removing first ten years apart}
plot(tset_phys_dist_mult, C_L.simu_phys_mult, col = "blue", lwd = 1, lty = 'solid', type = 'l', xlab = "Time (months)", ylab = "Proportional cover", xlim = c(0, max(tset_bio_dist_mult)), ylim = c(0, 1), main = "Structure-removing first")
lines(tset_phys_dist_mult, C_D.simu_phys_mult, col = "black", lwd = 1, lty = 'solid')
lines(tset_phys_dist_mult, M_S.simu_phys_mult, col = "green", lwd = 1, lty = 'solid')
lines(tset_phys_dist_mult, M_C.simu_phys_mult, col = "grey", lwd = 1, lty = 'solid')
lines(tset_bio_dist_mult, C_L.bio, col = "blue", lwd = 1, lty = 'solid')
lines(tset_bio_dist_mult, C_D.bio, col = "black", lwd = 1, lty = 'solid')
lines(tset_bio_dist_mult, M_S.bio, col = "green", lwd = 1, lty = 'solid')
lines(tset_bio_dist_mult, M_C.bio, col = "grey", lwd = 1, lty = 'solid')
abline(v=200, lty = 'dashed', col = "blue")
abline(v=320, lty = 'dashed', col = "red")
```

Structure retaining first, 10 years apart
```{r Structure-retaining first 10 years apart, echo = FALSE}

# Create time vector, holding vectors for each state variable, and set initial conditions
tset_bio_dist_mult <- seq(from = 0, to = 319)
C_L.bio <- NaN*tset_bio_dist_mult; C_L.bio[1] <- 0.6
C_D.bio <- NaN*tset_bio_dist_mult; C_D.bio[1] <- 0
M_S.bio <- NaN*tset_bio_dist_mult; M_S.bio[1] <- 0.05
M_C.bio <- NaN*tset_bio_dist_mult; M_C.bio[1] <- 0.05

# For() loop with if/else statement to simulate trajectories of spatial occupancy by live coral and macroalgae before and after disturbance
for(i in 2:length(tset_bio_dist_mult)){ # For every element after the first in the time vector...
  
  if(tset_bio_dist_mult[i] !=200){ # ...if the time point does not equal the time of disturbance, run the simulation as normal...

# calculate change in time  
  dt <- tset_bio_dist_mult[i] - tset_bio_dist_mult[i-1] 

  # Proportional cover at previous time step  
  C_L <- C_L.bio[i-1] 
  C_D <- C_D.bio[i-1]
  M_S <- M_S.bio[i-1]
  M_C <- M_C.bio[i-1]
 
# Change in proportional cover, converted to units of months (/5)
  dC_L <- (dt/5)*(g_C*C_L*(1-C_L-C_D-M_S-M_C) + gamma_C*(1-C_L-C_D-M_S-M_C) - g_M_C_L*C_L*(M_S + M_C) - d_C*C_L) 
  dC_D <- (dt/5)*(d_C*C_L + M_C*(h_M_C + d_M) - g_M*C_D*(M_S + M_C) - e*C_D)
  dM_S <- (dt/5)*(g_M*M_S*(1-C_L-C_D-M_S-M_C) + g_M*M_C*(1-C_L-C_D-M_S-M_C) + gamma_M*(1-C_L-C_D-M_S-M_C) - M_S*(h_M_S + d_M))
  dM_C <- (dt/5)*(g_M*M_C*C_D + g_M*M_S*C_D + g_M_C_L*C_L*(M_S + M_C) + gamma_M*C_D - M_C*(h_M_C + d_M + e))

# Proportional cover at new time step  
  C_L.bio[i] <- C_L.bio[i-1] + dC_L
  C_D.bio[i] <- C_D.bio[i-1] + dC_D
  M_S.bio[i] <- M_S.bio[i-1] + dM_S
  M_C.bio[i] <- M_C.bio[i-1] + dM_C
  
  }

  # ... but if the time step does equal the time of disturbance...
    else {
  
  dt <- tset_bio_dist_mult[i] - tset_bio_dist_mult[i-1]

  C_L <- C_L.bio[i-1]*0.5 # reduce live coral to 50% of its pre-disturbance level, and transfer this directly to the dead coral class
  C_D <- C_D.bio[i-1] + (C_L.bio[i-1]-C_L)
  M_S <- M_S.bio[i-1]
  M_C <- M_C.bio[i-1]
  
  dC_L <- (dt/5)*(g_C*C_L*(1-C_L-C_D-M_S-M_C) + gamma_C*(1-C_L-C_D-M_S-M_C) - g_M_C_L*C_L*(M_S + M_C) - d_C*C_L) 
  dC_D <- (dt/5)*(d_C*C_L + M_C*(h_M_C + d_M) - g_M*C_D*(M_S + M_C) - e*C_D)
  dM_S <- (dt/5)*(g_M*M_S*(1-C_L-C_D-M_S-M_C) + g_M*M_C*(1-C_L-C_D-M_S-M_C) + gamma_M*(1-C_L-C_D-M_S-M_C) - M_S*(h_M_S + d_M))
  dM_C <- (dt/5)*(g_M*M_C*C_D + g_M*M_S*C_D + g_M_C_L*C_L*(M_S + M_C) + gamma_M*C_D - M_C*(h_M_C + d_M + e))
  
  C_L.bio[i] <- C_L + dC_L
  C_D.bio[i] <- C_D + dC_D
  M_S.bio[i] <- M_S + dM_S
  M_C.bio[i] <- M_C + dM_C
  
  }
  
}
```


```{r Structure removing second 10 years apart, echo = FALSE}

# Create time vector, holding vectors for each state variable, and set initial conditions
tset_phys_dist_mult <- seq(from = 319, to = 1200)
C_L.simu_phys_mult <- NaN*tset_phys_dist_mult; C_L.simu_phys_mult[1] <- C_L.bio[320]
C_D.simu_phys_mult <- NaN*tset_phys_dist_mult; C_D.simu_phys_mult[1] <- C_D.bio[320]
M_S.simu_phys_mult <- NaN*tset_phys_dist_mult; M_S.simu_phys_mult[1] <- M_S.bio[320]
M_C.simu_phys_mult <- NaN*tset_phys_dist_mult; M_C.simu_phys_mult[1] <- M_C.bio[320]


# For() loop with if/else statement to simulate trajectories of spatial occupancy by live coral and macroalgae before and after disturbance

for(i in 2:length(tset_phys_dist_mult)){ # For every element after the first in the time vector...
  
  if(tset_phys_dist_mult[i] !=320){ # ...if the time point does not equal the time of disturbance, run the simulation as normal...
  
  dt <- tset_phys_dist_mult[i] - tset_phys_dist_mult[i-1] # calculate change in time  
 
  # Proportional cover at previous time step  
  C_L <- C_L.simu_phys_mult[i-1]
  C_D <- C_D.simu_phys_mult[i-1]
  M_S <- M_S.simu_phys_mult[i-1]
  M_C <- M_C.simu_phys_mult[i-1]
  
  # Change in proportional cover, converted to units of months (/5)
  dC_L <- (dt/5)*(g_C*C_L*(1-C_L-C_D-M_S-M_C) + gamma_C*(1-C_L-C_D-M_S-M_C) - g_M_C_L*C_L*(M_S + M_C) - d_C*C_L) 
  dC_D <- (dt/5)*(d_C*C_L + M_C*(h_M_C + d_M) - g_M*C_D*(M_S + M_C) - e*C_D)
  dM_S <- (dt/5)*(g_M*M_S*(1-C_L-C_D-M_S-M_C) + g_M*M_C*(1-C_L-C_D-M_S-M_C) + gamma_M*(1-C_L-C_D-M_S-M_C) - M_S*(h_M_S + d_M))
  dM_C <- (dt/5)*(g_M*M_C*C_D + g_M*M_S*C_D + g_M_C_L*C_L*(M_S + M_C) + gamma_M*C_D - M_C*(h_M_C + d_M + e))
 
  # Proportional cover at new time step   
  C_L.simu_phys_mult[i] <- C_L.simu_phys_mult[i-1] + dC_L
  C_D.simu_phys_mult[i] <- C_D.simu_phys_mult[i-1] + dC_D
  M_S.simu_phys_mult[i] <- M_S.simu_phys_mult[i-1] + dM_S
  M_C.simu_phys_mult[i] <- M_C.simu_phys_mult[i-1] + dM_C
  }
  
    else { # ... but if the time step does equal the time of disturbance...
  
      dt <- tset_phys_dist_mult[i] - tset_phys_dist_mult[i-1]

  # reduce all space holders to 50% of their pre-disturbance levels
  C_L <- C_L.simu_phys_mult[i-1]*0.5
  C_D <- C_D.simu_phys_mult[i-1]*0.5
  M_S <- M_S.simu_phys_mult[i-1]*0.5
  M_C <- M_C.simu_phys_mult[i-1]*0.5
  
  dC_L <- (dt/5)*(g_C*C_L*(1-C_L-C_D-M_S-M_C) + gamma_C*(1-C_L-C_D-M_S-M_C) - g_M_C_L*C_L*(M_S + M_C) - d_C*C_L) 
  dC_D <- (dt/5)*(d_C*C_L + M_C*(h_M_C + d_M) - g_M*C_D*(M_S + M_C) - e*C_D)
  dM_S <- (dt/5)*(g_M*M_S*(1-C_L-C_D-M_S-M_C) + g_M*M_C*(1-C_L-C_D-M_S-M_C) + gamma_M*(1-C_L-C_D-M_S-M_C) - M_S*(h_M_S + d_M))
  dM_C <- (dt/5)*(g_M*M_C*C_D + g_M*M_S*C_D + g_M_C_L*C_L*(M_S + M_C) + gamma_M*C_D - M_C*(h_M_C + d_M + e))
  
  C_L.simu_phys_mult[i] <- C_L+ dC_L
  C_D.simu_phys_mult[i] <- C_D+ dC_D
  M_S.simu_phys_mult[i] <- M_S+ dM_S
  M_C.simu_phys_mult[i] <- M_C+ dM_C
  }
  
}
```

Plot of time series for structure retaining followed by structure removing disturbance
```{r structure retaining first 10 years apart}
plot(tset_phys_dist_mult, C_L.simu_phys_mult, col = "blue", lwd = 1, lty = 'solid', type = 'l', xlab = "Time (months)", ylab = "Proportional cover", xlim = c(0, max(tset_phys_dist_mult)), ylim = c(0, 1), main = "Structure-retaining first")
lines(tset_phys_dist_mult, C_D.simu_phys_mult, col = "black", lwd = 1, lty = 'solid')
lines(tset_phys_dist_mult, M_S.simu_phys_mult, col = "green", lwd = 1, lty = 'solid')
lines(tset_phys_dist_mult, M_C.simu_phys_mult, col = "grey", lwd = 1, lty = 'solid')
lines(tset_bio_dist_mult, C_L.bio, col = "blue", lwd = 1, lty = 'solid')
lines(tset_bio_dist_mult, C_D.bio, col = "black", lwd = 1, lty = 'solid')
lines(tset_bio_dist_mult, M_S.bio, col = "green", lwd = 1, lty = 'solid')
lines(tset_bio_dist_mult, M_C.bio, col = "grey", lwd = 1, lty = 'solid')
abline(v=200, lty = 'dashed', col = "blue")
abline(v=320, lty = 'dashed', col = "red")
```

Structure retaining first, 2 years (The COTS-Oli scenario)
```{r Structure retaining first 2 years, echo = FALSE}

# Create time vector, holding vectors for each state variable, and set initial conditions
tset_bio_dist_mult <- seq(from = 0, to = 224)
C_L.bio <- NaN*tset_bio_dist_mult; C_L.bio[1] <- 0.6
C_D.bio <- NaN*tset_bio_dist_mult; C_D.bio[1] <- 0
M_S.bio <- NaN*tset_bio_dist_mult; M_S.bio[1] <- 0.05
M_C.bio <- NaN*tset_bio_dist_mult; M_C.bio[1] <- 0.05

# For() loop with if/else statement to simulate trajectories of spatial occupancy by live coral and macroalgae before and after disturbance
for(i in 2:length(tset_bio_dist_mult)){ # For every element after the first in the time vector...
  
  if(tset_bio_dist_mult[i] !=200){ # ...if the time point does not equal the time of disturbance, run the simulation as normal...

# calculate change in time  
  dt <- tset_bio_dist_mult[i] - tset_bio_dist_mult[i-1] 

  # Proportional cover at previous time step  
  C_L <- C_L.bio[i-1] 
  C_D <- C_D.bio[i-1]
  M_S <- M_S.bio[i-1]
  M_C <- M_C.bio[i-1]
 
# Change in proportional cover, converted to units of months (/5)
  dC_L <- (dt/5)*(g_C*C_L*(1-C_L-C_D-M_S-M_C) + gamma_C*(1-C_L-C_D-M_S-M_C) - g_M_C_L*C_L*(M_S + M_C) - d_C*C_L) 
  dC_D <- (dt/5)*(d_C*C_L + M_C*(h_M_C + d_M) - g_M*C_D*(M_S + M_C) - e*C_D)
  dM_S <- (dt/5)*(g_M*M_S*(1-C_L-C_D-M_S-M_C) + g_M*M_C*(1-C_L-C_D-M_S-M_C) + gamma_M*(1-C_L-C_D-M_S-M_C) - M_S*(h_M_S + d_M))
  dM_C <- (dt/5)*(g_M*M_C*C_D + g_M*M_S*C_D + g_M_C_L*C_L*(M_S + M_C) + gamma_M*C_D - M_C*(h_M_C + d_M + e))

# Proportional cover at new time step  
  C_L.bio[i] <- C_L.bio[i-1] + dC_L
  C_D.bio[i] <- C_D.bio[i-1] + dC_D
  M_S.bio[i] <- M_S.bio[i-1] + dM_S
  M_C.bio[i] <- M_C.bio[i-1] + dM_C
  
  }

  # ... but if the time step does equal the time of disturbance...
    else {
  
  dt <- tset_bio_dist_mult[i] - tset_bio_dist_mult[i-1]

  C_L <- C_L.bio[i-1]*0.1 # reduce live coral to 10% of its pre-disturbance level, and transfer this directly to the dead coral class
  C_D <- C_D.bio[i-1] + (C_L.bio[i-1]-C_L)
  M_S <- M_S.bio[i-1]
  M_C <- M_C.bio[i-1]
  
  dC_L <- (dt/5)*(g_C*C_L*(1-C_L-C_D-M_S-M_C) + gamma_C*(1-C_L-C_D-M_S-M_C) - g_M_C_L*C_L*(M_S + M_C) - d_C*C_L) 
  dC_D <- (dt/5)*(d_C*C_L + M_C*(h_M_C + d_M) - g_M*C_D*(M_S + M_C) - e*C_D)
  dM_S <- (dt/5)*(g_M*M_S*(1-C_L-C_D-M_S-M_C) + g_M*M_C*(1-C_L-C_D-M_S-M_C) + gamma_M*(1-C_L-C_D-M_S-M_C) - M_S*(h_M_S + d_M))
  dM_C <- (dt/5)*(g_M*M_C*C_D + g_M*M_S*C_D + g_M_C_L*C_L*(M_S + M_C) + gamma_M*C_D - M_C*(h_M_C + d_M + e))
  
  C_L.bio[i] <- C_L + dC_L
  C_D.bio[i] <- C_D + dC_D
  M_S.bio[i] <- M_S + dM_S
  M_C.bio[i] <- M_C + dM_C
  
  }
  
}
```


```{r Structure removing second 2 years, echo = FALSE}

# Create time vector, holding vectors for each state variable, and set initial conditions
tset_phys_dist_mult <- seq(from = 224, to = 1200)
C_L.simu_phys_mult <- NaN*tset_phys_dist_mult; C_L.simu_phys_mult[1] <- C_L.bio[224]
C_D.simu_phys_mult <- NaN*tset_phys_dist_mult; C_D.simu_phys_mult[1] <- C_D.bio[224]
M_S.simu_phys_mult <- NaN*tset_phys_dist_mult; M_S.simu_phys_mult[1] <- M_S.bio[224]
M_C.simu_phys_mult <- NaN*tset_phys_dist_mult; M_C.simu_phys_mult[1] <- M_C.bio[224]


# For() loop with if/else statement to simulate trajectories of spatial occupancy by live coral and macroalgae before and after disturbance

for(i in 2:length(tset_phys_dist_mult)){ # For every element after the first in the time vector...
  
  if(tset_phys_dist_mult[i] !=225){ # ...if the time point does not equal the time of disturbance, run the simulation as normal...
  
  dt <- tset_phys_dist_mult[i] - tset_phys_dist_mult[i-1] # calculate change in time  
 
  # Proportional cover at previous time step  
  C_L <- C_L.simu_phys_mult[i-1]
  C_D <- C_D.simu_phys_mult[i-1]
  M_S <- M_S.simu_phys_mult[i-1]
  M_C <- M_C.simu_phys_mult[i-1]
  
  # Change in proportional cover, converted to units of months (/5)
  dC_L <- (dt/5)*(g_C*C_L*(1-C_L-C_D-M_S-M_C) + gamma_C*(1-C_L-C_D-M_S-M_C) - g_M_C_L*C_L*(M_S + M_C) - d_C*C_L) 
  dC_D <- (dt/5)*(d_C*C_L + M_C*(h_M_C + d_M) - g_M*C_D*(M_S + M_C) - e*C_D)
  dM_S <- (dt/5)*(g_M*M_S*(1-C_L-C_D-M_S-M_C) + g_M*M_C*(1-C_L-C_D-M_S-M_C) + gamma_M*(1-C_L-C_D-M_S-M_C) - M_S*(h_M_S + d_M))
  dM_C <- (dt/5)*(g_M*M_C*C_D + g_M*M_S*C_D + g_M_C_L*C_L*(M_S + M_C) + gamma_M*C_D - M_C*(h_M_C + d_M + e))
 
  # Proportional cover at new time step   
  C_L.simu_phys_mult[i] <- C_L.simu_phys_mult[i-1] + dC_L
  C_D.simu_phys_mult[i] <- C_D.simu_phys_mult[i-1] + dC_D
  M_S.simu_phys_mult[i] <- M_S.simu_phys_mult[i-1] + dM_S
  M_C.simu_phys_mult[i] <- M_C.simu_phys_mult[i-1] + dM_C
  }
  
    else { # ... but if the time step does equal the time of disturbance...
  
      dt <- tset_phys_dist_mult[i] - tset_phys_dist_mult[i-1]

  # reduce all space holders to 50% of their pre-disturbance levels
  C_L <- C_L.simu_phys_mult[i-1]*0.1
  C_D <- C_D.simu_phys_mult[i-1]*0.1
  M_S <- M_S.simu_phys_mult[i-1]*0.1
  M_C <- M_C.simu_phys_mult[i-1]*0.1
  
  dC_L <- (dt/5)*(g_C*C_L*(1-C_L-C_D-M_S-M_C) + gamma_C*(1-C_L-C_D-M_S-M_C) - g_M_C_L*C_L*(M_S + M_C) - d_C*C_L) 
  dC_D <- (dt/5)*(d_C*C_L + M_C*(h_M_C + d_M) - g_M*C_D*(M_S + M_C) - e*C_D)
  dM_S <- (dt/5)*(g_M*M_S*(1-C_L-C_D-M_S-M_C) + g_M*M_C*(1-C_L-C_D-M_S-M_C) + gamma_M*(1-C_L-C_D-M_S-M_C) - M_S*(h_M_S + d_M))
  dM_C <- (dt/5)*(g_M*M_C*C_D + g_M*M_S*C_D + g_M_C_L*C_L*(M_S + M_C) + gamma_M*C_D - M_C*(h_M_C + d_M + e))
  
  C_L.simu_phys_mult[i] <- C_L+ dC_L
  C_D.simu_phys_mult[i] <- C_D+ dC_D
  M_S.simu_phys_mult[i] <- M_S+ dM_S
  M_C.simu_phys_mult[i] <- M_C+ dM_C
  }
  
}
```

Plot of time series for structure removing followed by structure retaining disturbance
```{r stucture retaining first two years}

plot(tset_phys_dist_mult, C_L.simu_phys_mult, col = "blue", lwd = 1, lty = 'solid', type = 'l', xlab = "Time (months)", ylab = "Proportional cover", xlim = c(0, max(tset_phys_dist_mult)), ylim = c(0, 1), main = "Structure-retaining first")
lines(tset_phys_dist_mult, C_D.simu_phys_mult, col = "black", lwd = 1, lty = 'solid')
lines(tset_phys_dist_mult, M_S.simu_phys_mult, col = "green", lwd = 1, lty = 'solid')
lines(tset_phys_dist_mult, M_C.simu_phys_mult, col = "grey", lwd = 1, lty = 'solid')
lines(tset_bio_dist_mult, C_L.bio, col = "blue", lwd = 1, lty = 'solid')
lines(tset_bio_dist_mult, C_D.bio, col = "black", lwd = 1, lty = 'solid')
lines(tset_bio_dist_mult, M_S.bio, col = "green", lwd = 1, lty = 'solid')
lines(tset_bio_dist_mult, M_C.bio, col = "grey", lwd = 1, lty = 'solid')
abline(v=200, lty = 'dashed', col = "blue")
abline(v=225, lty = 'dashed', col = "red")
```

Structure removing first, 2 years (The opposite of the COTS-Oli scenario)
```{r Structure-removing first 2 years, echo = FALSE}

# Create time vector, holding vectors for each state variable, and set initial conditions
tset_bio_dist_mult <- seq(from = 224, to = 1200)
C_L.bio <- NaN*tset_bio_dist_mult; C_L.bio[1] <- C_L.simu_phys_mult[224]
C_D.bio <- NaN*tset_bio_dist_mult; C_D.bio[1] <- C_D.simu_phys_mult[224]
M_S.bio <- NaN*tset_bio_dist_mult; M_S.bio[1] <- M_S.simu_phys_mult[224]
M_C.bio <- NaN*tset_bio_dist_mult; M_C.bio[1] <- M_C.simu_phys_mult[224]

# For() loop with if/else statement to simulate trajectories of spatial occupancy by live coral and macroalgae before and after disturbance
for(i in 2:length(tset_bio_dist_mult)){ # For every element after the first in the time vector...
  
  if(tset_bio_dist_mult[i] !=225){ # ...if the time point does not equal the time of disturbance, run the simulation as normal...

# calculate change in time  
  dt <- tset_bio_dist_mult[i] - tset_bio_dist_mult[i-1] 

  # Proportional cover at previous time step  
  C_L <- C_L.bio[i-1] 
  C_D <- C_D.bio[i-1]
  M_S <- M_S.bio[i-1]
  M_C <- M_C.bio[i-1]
 
# Change in proportional cover, converted to units of months (/5)
  dC_L <- (dt/5)*(g_C*C_L*(1-C_L-C_D-M_S-M_C) + gamma_C*(1-C_L-C_D-M_S-M_C) - g_M_C_L*C_L*(M_S + M_C) - d_C*C_L) 
  dC_D <- (dt/5)*(d_C*C_L + M_C*(h_M_C + d_M) - g_M*C_D*(M_S + M_C) - e*C_D)
  dM_S <- (dt/5)*(g_M*M_S*(1-C_L-C_D-M_S-M_C) + g_M*M_C*(1-C_L-C_D-M_S-M_C) + gamma_M*(1-C_L-C_D-M_S-M_C) - M_S*(h_M_S + d_M))
  dM_C <- (dt/5)*(g_M*M_C*C_D + g_M*M_S*C_D + g_M_C_L*C_L*(M_S + M_C) + gamma_M*C_D - M_C*(h_M_C + d_M + e))

# Proportional cover at new time step  
  C_L.bio[i] <- C_L.bio[i-1] + dC_L
  C_D.bio[i] <- C_D.bio[i-1] + dC_D
  M_S.bio[i] <- M_S.bio[i-1] + dM_S
  M_C.bio[i] <- M_C.bio[i-1] + dM_C
  
  }

  # ... but if the time step does equal the time of disturbance...
    else {
  
  dt <- tset_bio_dist_mult[i] - tset_bio_dist_mult[i-1]

  C_L <- C_L.bio[i-1]*0.1 # reduce live coral to 10% of its pre-disturbance level, and transfer this directly to the dead coral class
  C_D <- C_D.bio[i-1] + (C_L.bio[i-1]-C_L)
  M_S <- M_S.bio[i-1]
  M_C <- M_C.bio[i-1]
  
  dC_L <- (dt/5)*(g_C*C_L*(1-C_L-C_D-M_S-M_C) + gamma_C*(1-C_L-C_D-M_S-M_C) - g_M_C_L*C_L*(M_S + M_C) - d_C*C_L) 
  dC_D <- (dt/5)*(d_C*C_L + M_C*(h_M_C + d_M) - g_M*C_D*(M_S + M_C) - e*C_D)
  dM_S <- (dt/5)*(g_M*M_S*(1-C_L-C_D-M_S-M_C) + g_M*M_C*(1-C_L-C_D-M_S-M_C) + gamma_M*(1-C_L-C_D-M_S-M_C) - M_S*(h_M_S + d_M))
  dM_C <- (dt/5)*(g_M*M_C*C_D + g_M*M_S*C_D + g_M_C_L*C_L*(M_S + M_C) + gamma_M*C_D - M_C*(h_M_C + d_M + e))
  
  C_L.bio[i] <- C_L + dC_L
  C_D.bio[i] <- C_D + dC_D
  M_S.bio[i] <- M_S + dM_S
  M_C.bio[i] <- M_C + dM_C
  
  }
  
}
```


```{r Structure-retaining second 2 years, echo = FALSE}

# Create time vector, holding vectors for each state variable, and set initial conditions
tset_phys_dist_mult <- seq(from = 0, to = 224)
C_L.simu_phys_mult <- NaN*tset_phys_dist_mult; C_L.simu_phys_mult[1] <- 0.6
C_D.simu_phys_mult <- NaN*tset_phys_dist_mult; C_D.simu_phys_mult[1] <- 0
M_S.simu_phys_mult <- NaN*tset_phys_dist_mult; M_S.simu_phys_mult[1] <- 0.05
M_C.simu_phys_mult <- NaN*tset_phys_dist_mult; M_C.simu_phys_mult[1] <- 0.05


# For() loop with if/else statement to simulate trajectories of spatial occupancy by live coral and macroalgae before and after disturbance

for(i in 2:length(tset_phys_dist_mult)){ # For every element after the first in the time vector...
  
  if(tset_phys_dist_mult[i] !=200){ # ...if the time point does not equal the time of disturbance, run the simulation as normal...
  
  dt <- tset_phys_dist_mult[i] - tset_phys_dist_mult[i-1] # calculate change in time  
 
  # Proportional cover at previous time step  
  C_L <- C_L.simu_phys_mult[i-1]
  C_D <- C_D.simu_phys_mult[i-1]
  M_S <- M_S.simu_phys_mult[i-1]
  M_C <- M_C.simu_phys_mult[i-1]
  
  # Change in proportional cover, converted to units of months (/5)
  dC_L <- (dt/5)*(g_C*C_L*(1-C_L-C_D-M_S-M_C) + gamma_C*(1-C_L-C_D-M_S-M_C) - g_M_C_L*C_L*(M_S + M_C) - d_C*C_L) 
  dC_D <- (dt/5)*(d_C*C_L + M_C*(h_M_C + d_M) - g_M*C_D*(M_S + M_C) - e*C_D)
  dM_S <- (dt/5)*(g_M*M_S*(1-C_L-C_D-M_S-M_C) + g_M*M_C*(1-C_L-C_D-M_S-M_C) + gamma_M*(1-C_L-C_D-M_S-M_C) - M_S*(h_M_S + d_M))
  dM_C <- (dt/5)*(g_M*M_C*C_D + g_M*M_S*C_D + g_M_C_L*C_L*(M_S + M_C) + gamma_M*C_D - M_C*(h_M_C + d_M + e))
 
  # Proportional cover at new time step   
  C_L.simu_phys_mult[i] <- C_L.simu_phys_mult[i-1] + dC_L
  C_D.simu_phys_mult[i] <- C_D.simu_phys_mult[i-1] + dC_D
  M_S.simu_phys_mult[i] <- M_S.simu_phys_mult[i-1] + dM_S
  M_C.simu_phys_mult[i] <- M_C.simu_phys_mult[i-1] + dM_C
  }
  
    else { # ... but if the time step does equal the time of disturbance...
  
      dt <- tset_phys_dist_mult[i] - tset_phys_dist_mult[i-1]

  # reduce all space holders to 50% of their pre-disturbance levels
  C_L <- C_L.simu_phys_mult[i-1]*0.1
  C_D <- C_D.simu_phys_mult[i-1]*0.1
  M_S <- M_S.simu_phys_mult[i-1]*0.1
  M_C <- M_C.simu_phys_mult[i-1]*0.1
  
  dC_L <- (dt/5)*(g_C*C_L*(1-C_L-C_D-M_S-M_C) + gamma_C*(1-C_L-C_D-M_S-M_C) - g_M_C_L*C_L*(M_S + M_C) - d_C*C_L) 
  dC_D <- (dt/5)*(d_C*C_L + M_C*(h_M_C + d_M) - g_M*C_D*(M_S + M_C) - e*C_D)
  dM_S <- (dt/5)*(g_M*M_S*(1-C_L-C_D-M_S-M_C) + g_M*M_C*(1-C_L-C_D-M_S-M_C) + gamma_M*(1-C_L-C_D-M_S-M_C) - M_S*(h_M_S + d_M))
  dM_C <- (dt/5)*(g_M*M_C*C_D + g_M*M_S*C_D + g_M_C_L*C_L*(M_S + M_C) + gamma_M*C_D - M_C*(h_M_C + d_M + e))
  
  C_L.simu_phys_mult[i] <- C_L+ dC_L
  C_D.simu_phys_mult[i] <- C_D+ dC_D
  M_S.simu_phys_mult[i] <- M_S+ dM_S
  M_C.simu_phys_mult[i] <- M_C+ dM_C
  }
  
}
```

Plot of time series for structure removing followed by structure retaining disturbance
```{r structure-removing first two years}

plot(tset_phys_dist_mult, C_L.simu_phys_mult, col = "blue", lwd = 1, lty = 'solid', type = 'l', xlab = "Time (months)", ylab = "Proportional cover", xlim = c(0, max(tset_bio_dist_mult)), ylim = c(0, 1), main = "Structure-removing first")
lines(tset_phys_dist_mult, C_D.simu_phys_mult, col = "black", lwd = 1, lty = 'solid')
lines(tset_phys_dist_mult, M_S.simu_phys_mult, col = "green", lwd = 1, lty = 'solid')
lines(tset_phys_dist_mult, M_C.simu_phys_mult, col = "grey", lwd = 1, lty = 'solid')
lines(tset_bio_dist_mult, C_L.bio, col = "blue", lwd = 1, lty = 'solid')
lines(tset_bio_dist_mult, C_D.bio, col = "black", lwd = 1, lty = 'solid')
lines(tset_bio_dist_mult, M_S.bio, col = "green", lwd = 1, lty = 'solid')
lines(tset_bio_dist_mult, M_C.bio, col = "grey", lwd = 1, lty = 'solid')
abline(v=200, lty = 'dashed', col = "blue")
abline(v=225, lty = 'dashed', col = "red")
```